### 2021-02-26 proposal/proposal.tex
init proposal

\documentclass[11pt,conference]{IEEEtran}

% \IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.

\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage[colorlinks]{hyperref}
\usepackage[style=ieee]{biblatex}
\addbibresource{sources.bib}

\begin{document}

\title{Proposal: Proving Termination of \texttt{zelda-mosaic} Algorithms under Conditions on the Input}

\author{\IEEEauthorblockN{Justin Do}
\IEEEauthorblockA{\textit{Computer Science} \\
\textit{UNC Chapel Hill}\\
Chapel Hill, USA \\
\texttt{justindo@cs.unc.edu}}
\and
\IEEEauthorblockN{D. Ben Knoble}
\IEEEauthorblockA{\textit{Computer Science} \\
\textit{UNC Chapel Hill}\\
Chapel Hill, USA \\
\texttt{david3@cs.unc.edu}}
}

\maketitle

% \begin{abstract}
% \end{abstract}

% \begin{IEEEkeywords}
% \end{IEEEkeywords}

\section{Introduction}

We have previously developed \texttt{zelda-mosaic}~\cite{zelda_mosaic} which
includes MatLab~\cite{matlab} code to create tiled ``mosaics'' from a set of
smaller input images. We propose now to prove termination of our
mosaic-generation algorithms.

In the original work, we took a series of related images (specifically, from one
of the many ``Legend of Zelda'' ({\copyright} Nintendo) games) and automatically
stitched them together to form mosaics of game-related artwork.
{\figurename}~\ref{F:zelda-mosaic-sample} showcases an example output.

\begin{figure*}[!t]
    \centering
    \begin{subfigure}{0.35\textwidth}
        \includegraphics[width=\linewidth]{img/oracleofages-original.jpg}
        \caption{Original key-art}
    \end{subfigure}
    \begin{subfigure}{0.35\textwidth}
        \includegraphics[width=\linewidth]{img/oracle_of_ages_v1.png}
        \caption{Version 1 Mosaic}
    \end{subfigure}
    \begin{subfigure}{0.35\textwidth}
        \includegraphics[width=\linewidth]{img/oracle_of_ages_v2.png}
        \caption{Version 2 Mosaic}
    \end{subfigure}
    \begin{subfigure}{0.35\textwidth}
        \includegraphics[width=\linewidth]{img/oracle_of_ages_v3.png}
        \caption{Version 3 Mosaic}
    \end{subfigure}
    \begin{subfigure}{0.35\textwidth}
        \includegraphics[width=\linewidth]{img/oracle_of_ages_v1_smaller.png}
        \caption{Version 1 Mosaic with smaller inputs}
    \end{subfigure}
    \begin{subfigure}{0.35\textwidth}
        \includegraphics[width=\linewidth]{img/oracle_of_ages_v3_smaller.png}
        \caption{Version 3 Mosaic with smaller inputs}
    \end{subfigure}
    \caption{\texttt{zelda-mosaic} sample}\label{F:zelda-mosaic-sample}
\end{figure*}

We iteratively developed three versions after our initial proof-of-concept. Each
version refines the algorithm of the previous version (for details,
see~\ref{S:AlgDet}). Starting from a working prototype (version 1), we proceeded
to add limited duplication (version 2) and seam-blending (version 3). We also
further refined mosaic quality by stitching smaller images, at a performance
cost. The original presentation is publicly available~\cite{zelda_mosaic_pres},
as is a separate video-recording~\cite{zelda_mosaic_vid}.

Now we propose to examine these algorithms as implemented in the MatLab source
and prove, if possible, their termination given well-conditioned input. What
necessary conditions are remains to be seen, but likely includes constraints on
the sizes of each individual input image and the size of the key-art target
image.

\subsection{Motivation}

\section{Algorithm Details}\label{S:AlgDet}

\section{Related Work}

\section{Timeline}

\section{Contribution Report}

{\printbibliography}

\end{document}

### 2021-02-26 proposal/proposal.tex
correct casing

@@
 We have previously developed \texttt{zelda-mosaic}~\cite{zelda_mosaic} which
-includes MatLab~\cite{matlab} code to create tiled ``mosaics'' from a set of
+includes MATLAB~\cite{matlab} code to create tiled ``mosaics'' from a set of
 smaller input images. We propose now to prove termination of our
@@
 
-Now we propose to examine these algorithms as implemented in the MatLab source
+Now we propose to examine these algorithms as implemented in the MATLAB source
 and prove, if possible, their termination given well-conditioned input. What

### 2021-02-28 proposal/proposal.tex
incorporate/copy-edit justin's algorithms sections

@@
 \addbibresource{sources.bib}
+
+\renewcommand{\algorithmiccomment}[1]{\% #1}
 
@@
 stitched them together to form mosaics of game-related artwork.
-{\figurename}~\ref{F:zelda-mosaic-sample} showcases an example output.
+\figurename~\ref{F:zelda-mosaic-sample} showcases an example output. The program
+is extendable and has been successfully used on several different sizes of
+inputs and for applications beyond the original project.
 
@@
 version refines the algorithm of the previous version (for details,
-see~\ref{S:AlgDet}). Starting from a working prototype (version 1), we proceeded
-to add limited duplication (version 2) and seam-blending (version 3). We also
-further refined mosaic quality by stitching smaller images, at a performance
-cost. The original presentation is publicly available~\cite{zelda_mosaic_pres},
-as is a separate video-recording~\cite{zelda_mosaic_vid}.
+see~\ref{S:AlgDet}). Starting from a working prototype (version
+1,~\ref{S:algv1}), we proceeded to add limited duplication (version
+2,~\ref{S:algv2}) and seam-blending (version 3,~\ref{S:algv3}). We also further
+refined mosaic quality by stitching smaller images, at a performance cost. The
+original presentation is publicly available~\cite{zelda_mosaic_pres}, as is a
+separate video-recording~\cite{zelda_mosaic_vid}.
 
@@
 necessary conditions are remains to be seen, but likely includes constraints on
-the sizes of each individual input image and the size of the key-art target
+the sizes of each individual input image and the size of the key art target
 image.
@@
 
+% comments on proving properties of algorithms, MATLAB code, etc.
+
 \section{Algorithm Details}\label{S:AlgDet}
+
+In this section, we present the high-level details of the three algorithms about
+which we aim to prove termination. We begin with a brief overview of the
+relevant inputs before proceeding to give pseudo MATLAB-code for each version of
+the algorithm accompanied by textual explanation.
+
+\subsection{Background}
+
+Across each of the three versions, the inputs are the same. The three inputs are
+
+\begin{enumerate}
+    \item an image \textit{key art}, the target image the final mosaic will
+        replicate
+    \item a series of \textit{thumbnails}, smaller images which will form the
+        tiles of the mosaic
+    \item an integer \textit{size}, corresponding to the side-length of the
+        square thumbnails
+\end{enumerate}
+
+In the larger versions, we used a size of 75 pixels. In the smaller versions, we
+used 25 pixels. Each piece of key art is expected to have dimensions
+integer-multiples of the size input.
+
+The final output is a \textit{mosaic} of dimensions identical to the key art,
+where every \(size \times size\)-pixel subset is an image selected from the
+thumbnails.
+
+The MATLAB program reads these images, manipulates them as a \textit{cell
+array}\footnote{This is MATLAB's dynamically sized data-structure} of RGB
+values, and the result was written into a PNG file. Aside from general MATLAB
+control flow constructs, our code uses the built-in \textit{immse}~\cite{immse}
+function to calculate pixel-level differences between two images and the
+external \textit{MAT2TILES}~\cite{mat2tiles} library to break the key art images
+into tiles in a cell array format.
+
+\subsection{Algorithm for version 1}\label{S:algv1}
+
+Our first version of the algorithm is straightforward. It divides the key art
+into \(size \times size\)-pixel chunks. Each chunk is replaced with the
+corresponding ``best'' thumbnail. Here, ``best'' means the thumbnail with the
+lowest mean-squared error when compared to the original chunk. Details are
+reproduced in \figurename~\ref{alg:v1}.
+
+\begin{figure}[!t]
+    \textbf{Input}: image \(key\_art\), images \(thumbnails\), integer \(size\) \\
+    \textbf{Output}: image \(tiles\)
+    \begin{algorithmic}
+        \STATE{\(tiles \gets \textrm{MAT2TILES}(key\_art, size, size)\)}
+        \FOR{\(tile \in tiles\)}
+            \FOR{\(thumbnail, i \in thumbnails, [1..|thumbnails|]\)}
+                \STATE{\(mses_i \gets \textrm{immse}(tile, thumbnail)\)}
+            \ENDFOR
+            \STATE{\(best\_indices \gets \textrm{find} (mses = \min mses)\)}
+            \STATE{\(best\_index \gets best\_indices_0\)}
+            \STATE{\(tile \gets thumbnails_{best\_index}\)}
+        \ENDFOR
+    \RETURN{\(tiles\)}
+    \end{algorithmic}
+    \caption{Algorithm: Version 1}\label{alg:v1}
+\end{figure}
+
+\subsection{Algorithm for version 2}\label{S:algv2}
+
+The second version of the algorithm addresses a noticeable issue with our first
+attempt---namely, there is nothing preventing the algorithm from reusing the
+same thumbnail repeatedly, which can result in several recurrences of the same
+image in the final mosaic. This produces a boring result, and many of our
+version 1 outputs exhibit this repetition, in part because of the repeated use
+of similar color throughout the key art.
+
+This version of the algorithm keeps track of the thumbnails that have already
+been used, up to a threshold. It will search for the ``next-best'' option if the
+best thumbnails have already been used. Details are presented in
+\figurename~\ref{alg:v2}.
+
+The idea is to maintain a mapping from thumbnail to number of uses. When looking
+for the best thumbnail for a given chunk, only a thumbnail that has not exceeded
+the threshold in number of uses can be used. In effect, this limits each
+thumbnail to a certain (typically small) number of uses, preventing gratuitous
+repetition. This can also cause artifacts in the output, as we don't always get
+the most representative thumbnail for a given chunk.
+
+The threshold is chosen to allow thumbnails to be re-used enough to cover the
+entire mosaic, but not more than is absolutely required.
+
+\begin{figure}[!t]
+    \textbf{Input}: image \(key\_art\), images \(thumbnails\), integer \(size\) \\
+    \textbf{Output}: image \(tiles\)
+    \begin{algorithmic}
+        \STATE{\(tiles \gets \textrm{MAT2TILES}(key\_art, size, size)\)}
+        \STATE{\(used \gets \emptyset\)}
+        \STATE{\(threshold \gets 1\)}
+        \IF{\(|thumbnails| < |tiles|\)}
+            \STATE{\(threshold \gets \lceil{|tiles| / |thumbnails|}\rceil\)}
+        \ENDIF
+        \FOR{\(tile \in tiles\)}
+            \FOR{\(thumbnail, i \in thumbnails, [1..|thumbnails|]\)}
+                \STATE{\(mses_i \gets \textrm{immse}(tile, thumbnail)\)}
+            \ENDFOR
+            \STATE{\(best\_indices \gets \textrm{find} (mses = \min mses)\)}
+            \STATE{\(best\_index \gets best\_indices_0\)}
+            \WHILE{\(used_{best\_index} \ge threshold\)}
+                \STATE{\(mses_{best\_index} \gets \infty\)}
+                \STATE{\(best\_indices \gets \textrm{find} (mses = \min mses)\)}
+                \STATE{\(best\_index \gets best\_indices_0\)}
+            \ENDWHILE
+            \STATE{increment \(used_{best\_index}\)}
+            \STATE{\(tile \gets thumbnails_{best\_index}\)}
+        \ENDFOR
+        \RETURN{\(tiles\)}
+    \end{algorithmic}
+    \caption{Algorithm: Version 2}\label{alg:v2}
+\end{figure}
+
+\subsection{Algorithm for version 3}\label{S:algv3}
+
+The last noticeable issue we decided to tackle for version 3 was the seams
+between individual chunks. They aren't as noticeable in the outputs with smaller
+thumbnails, but in our original (\(75 \times 75\)) outputs there are visible
+grid lines where the chunks meet.
+
+To combat this, we tested a variety of blurring techniques to reduce the visual
+impact of the seams. In the process, we also aimed for producing a more coherent
+whole---i.e., a mosaic that look more like a true image than a tiled grid.
+
+We settled on combination of two techniques. The first technique is to blur the
+entire mosaic with a Gaussian filter (via the \textit{imgaussfilt} function),
+smoothing the image. The second technique is to blur each of the horizontal and
+vertical seams by replacing a narrow band around the seam with the average of
+the pixels in that region. Pseudocode is presented in \figurename~\ref{alg:v3}.
+
+\begin{figure}[!t]
+    \textbf{Input}: image \(key\_art\), images \(thumbnails\), integer \(size\) \\
+    \textbf{Output}: image \(tiles\)
+    \begin{algorithmic}
+        \STATE{}\COMMENT{Repeat Version 2}
+        \STATE{\(tiles \gets \textrm{Version2}(key\_art, thumbnails, size)\)}
+        \STATE{\(tiles \gets \textrm{imgaussfilt}(tiles)\)}
+        \STATE{\(strip\_size \gets \lfloor{size / 8}\rfloor\)}
+        \STATE{}\COMMENT{Blur by averaging}
+        \FOR{horizontal \(strip \in tiles\)}
+            \STATE{\(strip \gets \textrm{mean}({strip})\)}
+        \ENDFOR
+        \FOR{vertical \(strip \in tiles\)}
+            \STATE{\(strip \gets \textrm{mean}({strip})\)}
+        \ENDFOR
+        \RETURN{\(tiles\)}
+    \end{algorithmic}
+    \caption{Algorithm: Version 3}\label{alg:v3}
+\end{figure}
 

### 2021-02-28 proposal/proposal.tex
rm dead code
typo: look->looks
flesh out motivation

@@
 \documentclass[11pt,conference]{IEEEtran}
-
-% \IEEEoverridecommandlockouts
-% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
 
@@
 
-% comments on proving properties of algorithms, MATLAB code, etc.
+There are several factors that motivate our interest in this particular
+algorithm, its particular implementation, and its properties.
+
+Personally, it is a project we've already invested in, and one which we enjoy
+studying.
+
+Academically, it presents several unique opportunities. First, as far as we can
+tell, there have been no studies on either termination-properties of mosaic
+algorithms or of MATLAB code. Further, we have found no attempt to formalize
+details of the MATLAB programming language\footnote{This may in part be due to
+the proprietary, closed-source nature of MATLAB\@. Nevertheless, we hope to make
+an attempt, though it must be grounded in the knowledge that any proof can only
+hold if indeed the MATLAB environment faithfully implements the semantics it
+claims. There are formalizations of a particular MATLAB toolkit known as
+Stateflow~\cite{Hamon_2005,Hamon_2007} but none that we have found for general
+MATLAB programming.}. Thus already in studying this particularly narrow problem
+we have an opportunity to extend the space of studies in formal-methods and
+their applications to the programmer and their programs.
+
+Second, we as students of these formal methods lack experience in formalizing
+these details about specific algorithms. The Logical Foundations
+text~\cite{Pierce:SF1} has proven a great resource in learning about these
+principles on small examples, and our study of \texttt{zelda-mosaic} will make
+an excellent opportunity to extend our practical knowledge of the field.
+
+Lastly, termination and its decidability has always been a study of interest to
+the logician, the mathematician, and the computer scientist. From Hilbert's
+\textit{Entscheidungsproblem} to Turing's Halting problem~\cite{Cook_2011}, the
+world of algorithms is generally concerned with what is, indeed, computable.
+While we make no attempt here to prove termination of arbitrary or even
+restricted MATLAB programs (likely a Sisyphean task), we can at least attempt to
+push the boundaries of the undecidable Halting problem for this specific
+program.
 
@@
 impact of the seams. In the process, we also aimed for producing a more coherent
-whole---i.e., a mosaic that look more like a true image than a tiled grid.
+whole---i.e., a mosaic that looks more like a true image than a tiled grid.
 

### 2021-02-28 proposal/proposal.tex
abstract, related work, timeline, contributions

@@
 
-% \begin{abstract}
-% \end{abstract}
+\begin{abstract}
+    We propose to study a previous set of algorithms that we developed for
+    building mosaics, implemented in MATLAB\@. In particular, we are interested
+    in the formalization of the algorithms and the relevant MATLAB structures in
+    Coq and the termination of these algorithms under appropriate to-be-determined
+    conditions. We present details on the algorithms along with our motivations
+    for studying them. We then discuss an initial foray into related work.
+    Next, we propose a timeline for our progress. Lastly, we briefly identify
+    the major contributions of each author to this proposal.
+\end{abstract}
 
@@
 
+Because we are proving something about code that we have written ourselves,
+there is not a vast body of literature specific to our algorithms from which we
+can draw. There are however a number of related issues exposed by our code,
+which we will discuss here. In particular, the following are of interest and
+have been written about:
+
+\begin{itemize}
+    \item The correctness of MATLAB's \textit{cell array} data structure and its
+        representation in Coq;
+    \item The correctness of MATLAB's built-in numerical functions and matrix
+        operations; and
+    \item Proving termination of a particular program based on the inputs of the
+        program.
+\end{itemize}
+
+Since our program deals with a large amount of pixel information in memory as a
+matrix, it may be instructive to ensure that the representation of our data is
+robust and that assumptions made about them lead to the results that we expect.
+We are not the first group to demonstrate interest in proving properties about
+matrix representations. In the early days of Coq, a
+paper~\cite{magaud:hal-00955444} was published discussing the implementation of
+square matrices in Coq, as well as proofs involving basic operations such as
+addition and multiplication. This paper references \textit{Linear
+Algebra}~\cite{lin_alg}, a Coq library written by a colleague which formalizes
+a well-known linear algebra text~\cite{linear_algebra} and supports matrices with
+a differing number of rows and columns. It is not known at this time how these
+implementations compare with built-in Coq modules, but it will be useful to
+reference external matrix proofs and see how we might make similar arguments
+about matrix representations as they apply to our code.
+
+MATLAB is a closed-source piece of software, and little has been published about
+its correctness. In the official MATLAB forum, users have demonstrated interest
+in verifying the correctness of MATLAB algorithms~\cite{duenisch_2013} and the
+MATLAB staff reassure users that the algorithms have gone through rigorous
+testing, but this assurance does not carry the same weight as a formal
+verification. Indeed, as we have learned, these techniques fall on separate
+parts of a correctness spectrum, with testing demonstrating correctness less
+reliably than formal methods. The lack of related work in the MATLAB domain
+serves as further motivation to obtain results about the correctness of a subset
+of MATLAB features used in our code.
+
+Program termination is a topic of wide interest in the computing community, as
+it has both practical and of theoretical relevance given the famous halting
+problem. \citeauthor{Cook_2011}~\cite{Cook_2011} argue in their article that
+computer scientists are often hesitant to make claims about the termination of a
+particular program, but that it is often achievable. Their work draws from the
+field of compilers and makes the argument that one can prove the termination of
+a program given certain inputs, provided that there is sufficient analysis of
+the relation between the inputs and the possible paths that may be taken during
+execution\footnote{In many contexts, this involves things like symbolic
+execution (cf. Rosette~\cite{Rosette}).}. Many of the ideas in this work are
+formalized by \citeauthor{BLANQUI_2011} in the \textit{CoLoR} library for Coq,
+which is used primarily for proofs involving rewriting code and program
+termination~\cite{BLANQUI_2011}. These efforts demonstrate that the field is
+rich with material that relates termination and inputs, and, while we likely
+won't make significant contributions to this work, it will be interesting to see
+how well these arguments translate to the domain of scientific computing.
+
 \section{Timeline}
 
+Given the exploratory nature of this project, we may need to modify the
+properties we decide to prove and the scope of the project as we learn more
+about implementing our structures in Coq and decide the extent to which we can
+model elements of MATLAB code. Taking this into consideration, we estimate the
+following to be to be a reasonable timeline of project completion:
+
+\begin{itemize}
+    \item \textbf{03/01/21}: Project Proposal Due
+    \item \textbf{03/15/21}: Represent relevant data structures in Coq, as well
+        as primitive operations for matrix manipulation and accesses.
+    \item \textbf{03/29/21}: Scope of project finalized with formal problem
+        specification; sketch of theorems and lemmas that we expect to write in
+        the process of writing our proof.
+    \item \textbf{03/31/21}: Progress Report Due
+    \item \textbf{04/09/21}: Significant progress made on most theorems and
+        lemmas.
+    \item \textbf{04/19/21}: Most theorems and lemmas should be complete; change
+        scope of project where necessary.
+    \item \textbf{04/24/21}: All logic should be implemented; edits for readability should be the only changes henceforth.
+    \item \textbf{04/26/21}: Final Proof Due
+    \item \textbf{05/03/21}: Final Paper Due
+    \item \textbf{05/05/21}: Final Presentation Delivered
+\end{itemize}
+
 \section{Contribution Report}
+
+D. B. Knoble contributed the abstract, introduction, and motivation. He also
+contributed to the explanations of the algorithms. J. Do contributed the
+majority of the algorithm details, as well as the related work and timeline.
 

### 2021-04-09 progress-report/report.tex
progress report

\documentclass[11pt,conference]{IEEEtran}

\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage[inline]{enumitem}
\newlist{inlist}{enumerate*}{1}
\setlist[inlist]{label= (\alph*)}
\usepackage[colorlinks]{hyperref}
\hypersetup{%
    colorlinks=true,
    linkcolor=blue,
    urlcolor=cyan,
    citecolor=blue
}
\usepackage[style=ieee]{biblatex}
\addbibresource{sources.bib}

\newcommand{\matlab}{MATLAB}
\newcommand{\mmatlab}{\textmu\matlab}
\newcommand{\colorlib}{CoLoR}
\renewcommand{\algorithmiccomment}[1]{\% #1}

\begin{document}

\title{Progress Report: Proving Termination of \texttt{zelda-mosaic} Algorithms under Conditions on the Input}

\author{\IEEEauthorblockN{Justin Do}
\IEEEauthorblockA{\textit{Computer Science} \\
\textit{UNC Chapel Hill}\\
Chapel Hill, USA \\
\texttt{justindo@cs.unc.edu}}
\and
\IEEEauthorblockN{D. Ben Knoble}
\IEEEauthorblockA{\textit{Computer Science} \\
\textit{UNC Chapel Hill}\\
Chapel Hill, USA \\
\texttt{david3@cs.unc.edu}}
}

\maketitle

\begin{abstract}
    We report on a study of previous set of algorithms that we developed for
    building mosaics, implemented in \matlab\@. In particular, we are interested
    in the formalization of the algorithms and the relevant \matlab\ structures
    in Coq~\cite{Coq} and the termination of these algorithms under appropriate
    to-be-determined conditions. We present details on the algorithms along with
    our motivations for studying them. We then discuss an initial foray into
    related work. Next, we show an updated timeline for our progress. Lastly, we
    briefly identify the major contributions of each author to this proposal.
\end{abstract}

% \begin{IEEEkeywords}
% \end{IEEEkeywords}

\section{Summary of Proposal}

We have previously developed \texttt{zelda-mosaic}~\cite{zelda_mosaic} which
includes \matlab~\cite{matlab} code to create tiled ``mosaics'' from a set of
smaller input images. In our original proposal, we set out to prove the
termination of our mosaic-generation algorithms.

In the original work, we took a series of related images (specifically, from one
of the many ``Legend of Zelda'' ({\copyright} Nintendo) games) and automatically
stitched them together to form mosaics of game-related artwork.
\figurename~\ref{F:zelda-mosaic-sample} showcases an example output. The program
is extendable and has been successfully used on several different sizes of
inputs and for applications beyond the original project.

\begin{figure*}[!t]
    \centering
    \begin{subfigure}{0.35\textwidth}
        \includegraphics[width=\linewidth]{img/oracleofages-original.jpg}
        \caption{Original key-art}
    \end{subfigure}
    \begin{subfigure}{0.35\textwidth}
        \includegraphics[width=\linewidth]{img/oracle_of_ages_v1.png}
        \caption{Version 1 Mosaic}
    \end{subfigure}
    \begin{subfigure}{0.35\textwidth}
        \includegraphics[width=\linewidth]{img/oracle_of_ages_v2.png}
        \caption{Version 2 Mosaic}
    \end{subfigure}
    \begin{subfigure}{0.35\textwidth}
        \includegraphics[width=\linewidth]{img/oracle_of_ages_v3.png}
        \caption{Version 3 Mosaic}
    \end{subfigure}
    \begin{subfigure}{0.35\textwidth}
        \includegraphics[width=\linewidth]{img/oracle_of_ages_v1_smaller.png}
        \caption{Version 1 Mosaic with smaller inputs}
    \end{subfigure}
    \begin{subfigure}{0.35\textwidth}
        \includegraphics[width=\linewidth]{img/oracle_of_ages_v3_smaller.png}
        \caption{Version 3 Mosaic with smaller inputs}
    \end{subfigure}
    \caption{\texttt{zelda-mosaic} sample}\label{F:zelda-mosaic-sample}
\end{figure*}

We iteratively developed three versions after our initial proof-of-concept. Each
version refines the algorithm of the previous version (for details, the
pseudocode for each algorithm may be found in
\figurename{s}~\ref{alg:v1},~\ref{alg:v2}, and~\ref{alg:v3}). We also further
refined mosaic quality by stitching smaller images, at a performance cost. The
original presentation is publicly available~\cite{zelda_mosaic_pres}, as is a
separate video-recording~\cite{zelda_mosaic_vid}.

Our original goal was to examine these algorithms as implemented in the \matlab\
source and prove, if possible, their termination given well-conditioned input.
With some of the work now behind us, we have a better idea of what this project
entails and how we will be able to reach the finish line.

\section{Progress Report}

The starting point to our project was to fill in some knowledge gaps that we had
not covered in class prior to when we began. Since we planned to model a
program, we needed to know more about the modeling of programs and execution. To
this end, we completed work in \textit{Software Foundations} up to the chapter
\textit{Simple Imperative Programs}~\cite{Pierce:SF1}. This chapter provided
great insight on how to model statements and expressions, maintain state over a
program's execution, and define semantics. After completing this, we were better
able to outline the scope of the project.

Our progress thus far can be divided into two main categories:
\begin{inlist}
\item determining the subset of \matlab\ (dubbed \mmatlab) relevant to our program, and
\item creating data structures and functions suitable to model the behaviors of
    \mmatlab\@.
\end{inlist}

We also identify two main future tasks that will rely upon the completion of
these tasks:
\begin{inlist}
\item binding operations from the \mmatlab\ language to semantics in
    Coq~\cite{Coq}, and
\item using Hoare logic~\cite{Hoare_1969} to prove that the program terminates
    under what pre-conditions.
\end{inlist}

\subsection{\mmatlab}\label{S:mmatlab}

We give a brief overview on \mmatlab, the subset of \matlab\ that we model in
order to prove facts about our programs.

Aside from general programming patterns for control flow, the main syntax of
relevance in our program involves array/matrix accesses and updates, arithmetic
operations, and \matlab-specific primitives such as ranges. We have identified
all of the expressions and statements we expect to use and defined them as
inductive types, shown in \figurename~\ref{fig:inductivetypeexp} and
\figurename~\ref{fig:inductivetypestmt}.

\begin{figure}
    \centering
    \begin{align*}
        \mathsf{exp} \triangleq\;
        & |\; \mathsf{IntLiteral} \\
        & |\; \mathsf{FloatLiteral} \\
        & |\; \mathsf{AddExp} \\
        & |\; \mathsf{MultExp} \\
        & |\; \mathsf{DivExp} \\
        & |\; \mathsf{EqualsExp} \\
        & |\; \mathsf{NotEqualsExp} \\
        & |\; \mathsf{LTEqualsExp} \\
        & |\; \mathsf{Floor} \\
        & |\; \mathsf{VarRefExp} \\
        & |\; \mathsf{RangeExp} \\
        & |\; \mathsf{MatrixLiteral} \\
        & |\; \mathsf{IndexExp} \\
        & |\; \mathsf{SqueezeExpr} \\
        & |\; \mathsf{SizeExpr} \\
        & |\; \mathsf{LengthExpr} \\
        & |\; \mathsf{ProdExpr} \\
        & |\; \mathsf{SumExpr} \\
        & |\; \mathsf{MinExpr} \\
        & |\; \mathsf{MaxExpr} \\
        & |\; \mathsf{PointWiseExpExp} \\
        & |\; \mathsf{ZerosExp} \\
        & |\; \mathsf{OnesExp} \\
        & |\; \mathsf{InfExp} \\
        & |\; \mathsf{FindExp}
    \end{align*}
    \caption{Skeleton of abstract syntax for expressions used in
    \mmatlab\@.}\label{fig:inductivetypeexp}
\end{figure}

\begin{figure}
    \centering
    \begin{align*}
        \mathsf{statement} \triangleq\;
        & |\; \mathsf{NoOp} \\
        & |\; \mathsf{Error} \\
        & |\; \mathsf{ExprS} \\
        & |\; \mathsf{AssignS} \\
        & |\; \mathsf{IndexedAssignS} \\
        & |\; \mathsf{SeqS} \\
        & |\; \mathsf{WhileS} \\
        & |\; \mathsf{IfThenElseS} \\
    \end{align*}
    \caption{Skeleton of abstract syntax for statements used in
    \mmatlab\@.}\label{fig:inductivetypestmt}
\end{figure}

Much of the abstract syntax is relatively self-explanatory, so complete
definitions are elided here. However, we wish to draw attention to some of the
less apparent and more interesting syntax present in \mmatlab\@:

\begin{itemize}

    \item \textsf{AddExp}: While this is nominally a regular addition operator,
        \matlab's addition is interesting in that it has overloaded parameter
        types and expected behavior for each. The addition of two numbers is a
        number and the addition of two matrices results in element-wise
        addition, as one might expect. However, the sum of a number and a
        matrix, is a matrix with all elements incremented by the number. Similar
        behaviors hold for \textsf{MultExp} and \textsf{DivExp}.

    \item \textsf{RangeExp}: Like Python, \matlab\ has an expression where a
        starting number and stopping number may be provided, generating an
        inclusive list of all numbers between them.

    \item \textsf{SqueezeExpr}: For multi-dimensional arrays, sometimes the size
        of a particular dimension will be 1. \texttt{squeeze} is a function that
        reduces dimensionality by discarding or ``squeezing'' these superfluous
        dimensions.

\end{itemize}

\subsection{Data Structures and Functions}

Our program is used for image processing; thus, many computations are made on
images, which are represented as arrays in \matlab\@. Additionally, we deal with
matrices of images, so we need to model multi-dimensional matrices. Since
multi-dimensional matrices can have arbitrarily many dimensions, we decided the
best abstraction for this in Coq would be built on top of the \textsf{Vector}
type which was originally implemented in the \colorlib\
library~\cite{BLANQUI_2011} and eventually ported to the Coq standard library.
In Coq, \textsf{Vector} is a dependent type~\cite{Bove2009,Thorsten_2010}; the
number of elements in a \textsf{Vector} is part of its type, written
\texttt{Vector.t A n} for a vector containing \(n\) elements of type \(A\).
Since \textsf{Vector}s are polymorphic, we noted that a \textsf{Vector} could
contain \textsf{Vector}s provided they are all the same type (which includes
length, as noted). That is, we can stack \textsf{Vector}s to form, \emph{e.g.},
\texttt{Vector.t (Vector.t A m) n}. From there, we defined a type-constructor
\textsf{Matrix} taking a list of dimensions, which represents a \textsf{Vector}
of \textsf{Vector}s. It can be used as the type of an arbitrary
\(n\)-dimensional matrix containing elements of type \(A\).

Due to the implementation of \textsf{Vector} in the Coq standard library, we
needed to greatly extend the functionality of our \textsf{Matrix} type. To that
end, we have begun implementing \textsf{Matrix} functions such as element
access, which is non-trivial for matrices with an arbitrary number of
dimensions\footnote{In particular because of the dependent types; see also
\url{https://stackoverflow.com/q/66863226/4400820}}. Additionally, the existing
implementation of element access in Coq's \textsf{Vector} type uses the
unfamiliar \textsf{Fin} type, an abstraction for carrying proofs that a
particular number belongs to finite range of numbers. In order to access an
element of a \textsf{Vector}, one must supply such a proof; for our purposes, we
use \texttt{Fin.of\_nat}, which gives either such a proof or a witness that the
number is outside the range. This behavior is captured by the \textsf{sumbool}
built-in, which required further research on our behalf. Given the number of
ways that matrices can be indexed in \matlab\ and the complexity introduced by
Coq's implementation of \textsf{Vector}s, we see this as the main bottleneck for
our progress. We still need to implement many \textsf{Matrix} operations, such
as accesses over ranges and over an entire dimension.

\subsection{Binding Operations and Semantics}

While we have not yet bound any of our functions or data types to the statements
or expressions in \S~\ref{S:mmatlab}, we have naturally been thinking about them
throughout the whole process. The functions we write are intended to implement
the evaluation of the corresponding expressions. In the style of
\texttt{Imp}~\cite{Pierce:SF1}, we intend to write a deterministic expression
evaluation function; we will then provide semantics for program execution.
Execution cannot be defined as a function in Coq because it may be
non-terminating. We have had to consider what operations are be allowed and
whether checks for undesired behavior (\emph{e.g.}, dynamic type mismatches)
should be implemented in the underlying functions or in a semantic check.

\subsection{Hoare Logic}

We have a significant amount of work to do in modeling \mmatlab, but the
blue-sky goal remains to prove the termination of our program given that certain
preconditions hold. Ben has been studying Hoare logic~\cite{Hoare_1969}, the
relevant background material for this problem, as part of his Thesis
requirement. What will be possible to prove will be determined in part by what
we are able to successfully model, given the time constraints. However, as a
preliminary step, we have studied partial Hoare logic and total Hoare logic to
provide more context for the task ahead of us. In this exploration, we found
that partial Hoare logic is a weak statement about programs; namely, given
preconditions \(P\) and postconditions \(Q\) and a program \(C\), the Hoare
triple \(\{P\}C\{Q\}\) asserts that \emph{if} \(C\) terminates, then the
pre-conditions imply the post-conditions. Total Hoare logic, by contrast, is a
stronger version of this claim: \(C\) is asserted to terminate. Thus a proof in
total Hoare logic necessitates a proof that \(C\) terminates; the inference
rules are accordingly adjusted. The primary adjustment is to the rule for while
loops, which are guaranteed to terminate by relation to a decreasing sequence
(with respect to a well-founded order). Given that we are trying to prove
termination, it seems likely that we will need to use total Hoare logic in the
relevant proofs.

\begin{figure}[!t]
    \textbf{Input}: image \(key\_art\), images \(thumbnails\), integer \(size\) \\
    \textbf{Output}: image \(tiles\)
    \begin{algorithmic}
        \STATE{\(tiles \gets \textrm{MAT2TILES}(key\_art, size, size)\)}
        \FOR{\(tile \in tiles\)}
            \FOR{\(thumbnail, i \in thumbnails, [1..|thumbnails|]\)}
                \STATE{\(mses_i \gets \textrm{immse}(tile, thumbnail)\)}
            \ENDFOR
            \STATE{\(best\_indices \gets \textrm{find} (mses = \min mses)\)}
            \STATE{\(best\_index \gets best\_indices_0\)}
            \STATE{\(tile \gets thumbnails_{best\_index}\)}
        \ENDFOR
    \RETURN{\(tiles\)}
    \end{algorithmic}
    \caption{Algorithm: Version 1}\label{alg:v1}
\end{figure}

\begin{figure}[!t]
    \textbf{Input}: image \(key\_art\), images \(thumbnails\), integer \(size\) \\
    \textbf{Output}: image \(tiles\)
    \begin{algorithmic}
        \STATE{\(tiles \gets \textrm{MAT2TILES}(key\_art, size, size)\)}
        \STATE{\(used \gets \emptyset\)}
        \STATE{\(threshold \gets 1\)}
        \IF{\(|thumbnails| < |tiles|\)}
            \STATE{\(threshold \gets \lceil{|tiles| / |thumbnails|}\rceil\)}
        \ENDIF
        \FOR{\(tile \in tiles\)}
            \FOR{\(thumbnail, i \in thumbnails, [1..|thumbnails|]\)}
                \STATE{\(mses_i \gets \textrm{immse}(tile, thumbnail)\)}
            \ENDFOR
            \STATE{\(best\_indices \gets \textrm{find} (mses = \min mses)\)}
            \STATE{\(best\_index \gets best\_indices_0\)}
            \WHILE{\(used_{best\_index} \ge threshold\)}
                \STATE{\(mses_{best\_index} \gets \infty\)}
                \STATE{\(best\_indices \gets \textrm{find} (mses = \min mses)\)}
                \STATE{\(best\_index \gets best\_indices_0\)}
            \ENDWHILE
            \STATE{increment \(used_{best\_index}\)}
            \STATE{\(tile \gets thumbnails_{best\_index}\)}
        \ENDFOR
        \RETURN{\(tiles\)}
    \end{algorithmic}
    \caption{Algorithm: Version 2}\label{alg:v2}
\end{figure}

\begin{figure}[!t]
    \textbf{Input}: image \(key\_art\), images \(thumbnails\), integer \(size\) \\
    \textbf{Output}: image \(tiles\)
    \begin{algorithmic}
        \STATE{}\COMMENT{Repeat Version 2}
        \STATE{\(tiles \gets \textrm{Version2}(key\_art, thumbnails, size)\)}
        \STATE{\(tiles \gets \textrm{imgaussfilt}(tiles)\)}
        \STATE{\(strip\_size \gets \lfloor{size / 8}\rfloor\)}
        \STATE{}\COMMENT{Blur by averaging}
        \FOR{horizontal \(strip \in tiles\)}
            \STATE{\(strip \gets \textrm{mean}({strip})\)}
        \ENDFOR
        \FOR{vertical \(strip \in tiles\)}
            \STATE{\(strip \gets \textrm{mean}({strip})\)}
        \ENDFOR
        \RETURN{\(tiles\)}
    \end{algorithmic}
    \caption{Algorithm: Version 3}\label{alg:v3}
\end{figure}

\section{Goals and Timeline}

As we neglected to include them in our original proposal, we now propose the
following goals, taking into consideration the progress that we have made so far
with the project:

\begin{enumerate}

    \item \textbf{Minimum Viable Product}: Successful modelling of a majority of
        the desired features corresponding to matrix operations in \mmatlab\@.
        Given the lack of multi-dimensional matrix libraries in Coq and the
        necessary additions of several features in the Coq standard library,
        this alone would be a notable achievement for us.

    \item \textbf{Desired Goal}: Successful binding of our matrix operations to
        \mmatlab\ statements and expressions, including the ability to evaluate
        programs written in \mmatlab\@. A proof of termination, or sketch of
        one, for our V1 algorithm is provided.

    \item \textbf{Stretch Goal}: All of the above, and the addition of all
        operations used in V2 and V3 algorithms (most notably, for loops that
        increment across a range at varying intervals). Proofs that use
        these features.

\end{enumerate}

Given what we have encountered so far, we propose the following modified
timeline for project completion.

\begin{itemize}
    \item \textbf{03/31/21}: Progress Report Due
    \item \textbf{04/09/21}: Significant progress made on most matrix operations
        and data structures.
    \item \textbf{04/16/21}: Binding \mmatlab\ syntax to operations written
        in Coq; definition of semantics and program evaluator; formal proof
        specification written.
    \item \textbf{04/24/21}: Proof(s) should be written or scope adjusted
        accordingly. All logic should be implemented; edits for readability
        should be the main changes henceforth.
    \item \textbf{04/26/21}: Final Proof Due
    \item \textbf{05/03/21}: Final Paper Due
    \item \textbf{05/05/21}: Final Presentation Delivered
\end{itemize}

We provide the following ``diff'' between the scope of our original project
proposal and our new proposal:
\begin{itemize}
    \item Heightened emphasis on implementation of matrix operations, since they
        are proving to be non-trivial.
    \item Addition of goal tiers.
    \item Discussion on Hoare logic.
\end{itemize}

{\printbibliography}

\end{document}

### 2021-04-29 paper/paper.tex
add paper skeleton
- copy report
- add ability to input coq-tex files
- add SO to sources
paper: kill goals/timeline section

--- 2021-04-09 progress-report/report.tex
@@
 \renewcommand{\algorithmiccomment}[1]{\% #1}
+\newcommand{\inputcoq}[1]{\InputIfFileExists{#1.v_tex}{}{\typeout{No file #1.v_tex.}}}
 
@@
 
-\title{Progress Report: Proving Termination of \texttt{zelda-mosaic} Algorithms under Conditions on the Input}
+\title{Final paper: Proving Termination of \texttt{zelda-mosaic} Algorithms under Conditions on the Input}
 
@@
 
-\section{Goals and Timeline}
-
-As we neglected to include them in our original proposal, we now propose the
-following goals, taking into consideration the progress that we have made so far
-with the project:
-
-\begin{enumerate}
-
-    \item \textbf{Minimum Viable Product}: Successful modelling of a majority of
-        the desired features corresponding to matrix operations in \mmatlab\@.
-        Given the lack of multi-dimensional matrix libraries in Coq and the
-        necessary additions of several features in the Coq standard library,
-        this alone would be a notable achievement for us.
-
-    \item \textbf{Desired Goal}: Successful binding of our matrix operations to
-        \mmatlab\ statements and expressions, including the ability to evaluate
-        programs written in \mmatlab\@. A proof of termination, or sketch of
-        one, for our V1 algorithm is provided.
-
-    \item \textbf{Stretch Goal}: All of the above, and the addition of all
-        operations used in V2 and V3 algorithms (most notably, for loops that
-        increment across a range at varying intervals). Proofs that use
-        these features.
-
-\end{enumerate}
-
-Given what we have encountered so far, we propose the following modified
-timeline for project completion.
-
-\begin{itemize}
-    \item \textbf{03/31/21}: Progress Report Due
-    \item \textbf{04/09/21}: Significant progress made on most matrix operations
-        and data structures.
-    \item \textbf{04/16/21}: Binding \mmatlab\ syntax to operations written
-        in Coq; definition of semantics and program evaluator; formal proof
-        specification written.
-    \item \textbf{04/24/21}: Proof(s) should be written or scope adjusted
-        accordingly. All logic should be implemented; edits for readability
-        should be the main changes henceforth.
-    \item \textbf{04/26/21}: Final Proof Due
-    \item \textbf{05/03/21}: Final Paper Due
-    \item \textbf{05/05/21}: Final Presentation Delivered
-\end{itemize}
-
-We provide the following ``diff'' between the scope of our original project
-proposal and our new proposal:
-\begin{itemize}
-    \item Heightened emphasis on implementation of matrix operations, since they
-        are proving to be non-trivial.
-    \item Addition of goal tiers.
-    \item Discussion on Hoare logic.
-\end{itemize}
-
 {\printbibliography}

### 2021-05-02 paper/paper.tex
paper: fix title
paper: fix figure locations
paper: use func/var macros (esp. useful later)

@@
 \newcommand{\inputcoq}[1]{\InputIfFileExists{#1.v_tex}{}{\typeout{No file #1.v_tex.}}}
+\newcommand{\var}[1]{\mathit{#1}}
+\newcommand{\func}[1]{\mathsf{#1}}
 
@@
 
-\title{Final paper: Proving Termination of \texttt{zelda-mosaic} Algorithms under Conditions on the Input}
+\title{Final paper: Towards Proving Termination of \texttt{zelda-mosaic} Algorithms under Conditions on the Input}
 
@@
 
-\begin{figure*}[!t]
+\begin{figure*}[pht]
     \centering
@@
 
-\begin{figure}
+\begin{figure}[ht]
     \centering
     \begin{align*}
-        \mathsf{exp} \triangleq\;
-        & |\; \mathsf{IntLiteral} \\
-        & |\; \mathsf{FloatLiteral} \\
-        & |\; \mathsf{AddExp} \\
-        & |\; \mathsf{MultExp} \\
-        & |\; \mathsf{DivExp} \\
-        & |\; \mathsf{EqualsExp} \\
-        & |\; \mathsf{NotEqualsExp} \\
-        & |\; \mathsf{LTEqualsExp} \\
-        & |\; \mathsf{Floor} \\
-        & |\; \mathsf{VarRefExp} \\
-        & |\; \mathsf{RangeExp} \\
-        & |\; \mathsf{MatrixLiteral} \\
-        & |\; \mathsf{IndexExp} \\
-        & |\; \mathsf{SqueezeExpr} \\
-        & |\; \mathsf{SizeExpr} \\
-        & |\; \mathsf{LengthExpr} \\
-        & |\; \mathsf{ProdExpr} \\
-        & |\; \mathsf{SumExpr} \\
-        & |\; \mathsf{MinExpr} \\
-        & |\; \mathsf{MaxExpr} \\
-        & |\; \mathsf{PointWiseExpExp} \\
-        & |\; \mathsf{ZerosExp} \\
-        & |\; \mathsf{OnesExp} \\
-        & |\; \mathsf{InfExp} \\
-        & |\; \mathsf{FindExp}
+        \func{exp} \triangleq\;
+        & |\; \func{IntLiteral} \\
+        & |\; \func{FloatLiteral} \\
+        & |\; \func{AddExp} \\
+        & |\; \func{MultExp} \\
+        & |\; \func{DivExp} \\
+        & |\; \func{EqualsExp} \\
+        & |\; \func{NotEqualsExp} \\
+        & |\; \func{LTEqualsExp} \\
+        & |\; \func{Floor} \\
+        & |\; \func{VarRefExp} \\
+        & |\; \func{RangeExp} \\
+        & |\; \func{MatrixLiteral} \\
+        & |\; \func{IndexExp} \\
+        & |\; \func{SqueezeExpr} \\
+        & |\; \func{SizeExpr} \\
+        & |\; \func{LengthExpr} \\
+        & |\; \func{ProdExpr} \\
+        & |\; \func{SumExpr} \\
+        & |\; \func{MinExpr} \\
+        & |\; \func{MaxExpr} \\
+        & |\; \func{PointWiseExpExp} \\
+        & |\; \func{ZerosExp} \\
+        & |\; \func{OnesExp} \\
+        & |\; \func{InfExp} \\
+        & |\; \func{FindExp}
     \end{align*}
@@
 
-\begin{figure}
+\begin{figure}[ht]
     \centering
     \begin{align*}
-        \mathsf{statement} \triangleq\;
-        & |\; \mathsf{NoOp} \\
-        & |\; \mathsf{Error} \\
-        & |\; \mathsf{ExprS} \\
-        & |\; \mathsf{AssignS} \\
-        & |\; \mathsf{IndexedAssignS} \\
-        & |\; \mathsf{SeqS} \\
-        & |\; \mathsf{WhileS} \\
-        & |\; \mathsf{IfThenElseS} \\
+        \func{statement} \triangleq\;
+        & |\; \func{NoOp} \\
+        & |\; \func{Error} \\
+        & |\; \func{ExprS} \\
+        & |\; \func{AssignS} \\
+        & |\; \func{IndexedAssignS} \\
+        & |\; \func{SeqS} \\
+        & |\; \func{WhileS} \\
+        & |\; \func{IfThenElseS} \\
     \end{align*}
@@
 
-\begin{figure}[!t]
+\begin{figure}[ht]
     \textbf{Input}: image \(key\_art\), images \(thumbnails\), integer \(size\) \\

### 2021-05-02 paper/paper.tex
paper: draft results part A

@@
 \usepackage{algorithmic}
+\usepackage{array,booktabs}
 \usepackage{graphicx}
@@
 \setlist[inlist]{label= (\alph*)}
+\usepackage{mathpartir,semantic}
 \usepackage[colorlinks]{hyperref}
@@
 \newcommand{\func}[1]{\mathsf{#1}}
+\newcommand{\iname}[1]{\textsf{#1}}
 
@@
 
+\section{Results}\label{S:results}
+
+In this section, we present the results of our work towards proving termination
+of our mosaic algorithm. As noted, our primary accomplishment is the definition
+of a workable matrix type and operations on it. This matrix type is intended to
+support the operations needed to implement expression evaluation for \mmatlab, a
+crucial component of implementing the language's semantics. It remains to finish
+implementing expression evaluation and the necessary matrix operations, to give
+operation semantics for the statements of \mmatlab, to give a total Hoare logic
+for the language, and then finally to give the proofs of termination of the
+algorithms found in \figurename{s}~\ref{alg:v1},~\ref{alg:v2}, and~\ref{alg:v3}.
+
+We present, in order,
+\begin{inlist}
+\item the Coq definition of the polymorphic matrix type \texttt{matrix A} and
+    corresponding operations;
+\item key theorems and sketches of their proofs about the matrix type and its
+    operations; and
+\item advanced tactics, automation, and ideas for Coq proofs used in the
+    development.
+\end{inlist}
+
+% TODO: update zelda_mosaic_proof version with final commit
+
+All code may be found at the project repository~\cite{zelda_mosaic_proof}. Most
+of the work in this section corresponds to the file \texttt{Lib Matrix.v}.
+
+\subsection{Defining Matrices}\label{S:matrix_defn}
+
+First, we define the matrix type. After noting that it allows invalid matrices
+to be formed, we define the well-formed property on matrices in two equivalent
+ways. Finally, we give an overview of the Coq definitions of three operations on
+matrices: computing their shape, linearizing, and linear indexing. We briefly
+mention the beginning of our work on multi-dimensional and range-based indexing.
+
+\subsubsection{The matrix type}
+
+As mentioned in \S~\ref{S:design_dependent}, our original formulation of
+matrices as nested dependent \texttt{Vector.t}s quickly grew in complexity.
+We abandoned that approach in favor of a type that is not guaranteed to behave
+well, but whose definition makes defining operations more
+natural~\cite{SO_2021_2}. This approach removes the proofs that dependent types
+provide automatically, so we have to give those ourselves.
+
+The new design pairs the dimensionality of a matrix, called its \emph{shape},
+with its \emph{contents}. The pairing is done via \emph{Record} types (see the
+Coq manual~\cite{Coq}), which are tuples with named elements. The definition
+follows below:
+
+\begin{align*}
+    & \func{matrix} (\var{A}: \var{Type}) \triangleq \\
+    & \{
+        \func{shape}: \func{list}\, \mathbb{N},
+        \func{contents}: \func{matrix\_content}\, \var{A}
+    \}
+\end{align*}
+
+The shape is a list of naturals. Each natural corresponds (in a well-formed
+matrix) to the number of elements of the matrix in that dimension. For example,
+a vector of five (5) elements has shape \texttt{[5]}, while a two-dimensional
+matrix of three-by-two elements has shape \texttt{[3; 2]}. The vacuous case,
+when the shape is the empty list, is a single scalar value: there is no matrix.
+
+In parallel, the contents are wrapped in an inductive polymorphic type
+\(\func{matrix\_content}\, \var{A}\). The type's two constructors correspond to
+the two cases above: a \(\func{Matrix}\) contains a list of
+\(\func{matrix\_content}\, \var{A}\) values. The \(\func{Matrix}\) constructor,
+in a well-formed matrix, corresponds to one dimension of the shape. A
+\(\func{Scalar}\) contains exactly a value of type \(\var{A}\) and is the
+conceptually the base-case for the matrix type.
+
+\begin{align*}
+    & \func{matrix\_content} (\var{A}: \var{Type}) \triangleq\; \\
+    & |\; \func{Scalar}: \var{A} \to \func{matrix\_content}\, \var{A} \\
+    & |\; \func{Matrix}: \func{list}\, (\func{matrix\_content}\, \var{A}) \to \func{matrix\_content}\, \var{A}
+\end{align*}
+
+We provide a few concrete examples of mathematical objects in the space of
+scalars and multi-dimensional matrices containing a single type of object in
+Table~\ref{T:matrices}. Each example shows standard mathematical notation, the
+Coq notation using our matrix type, and the resulting type of the value in Coq.
+
+\begin{table*}[ht]
+    \centering
+    \begin{tabular}{c >{\ttfamily}c >{\ttfamily}c}
+        Mathematical notation & Coq notation & Coq type \\ \toprule
+        \(5\) & \{| shape := []; contents := Scalar 5 |\} & matrix nat \\
+        \(\begin{bmatrix}
+            1 & 2 & 3
+        \end{bmatrix}\)
+        & \{| shape := [3]; contents := Matrix [Scalar 1; Scalar 2; Scalar 3] |\} & matrix nat \\
+        \\
+        \(\begin{bmatrix}
+            1 & 2 \\
+            3 & 4 \\
+            5 & 6
+        \end{bmatrix}\)
+        & \{| shape := [3; 2]; contents := Matrix [
+        & matrix nat \\
+        & Matrix [Scalar 1; Scalar 2]; & \\
+        & Matrix [Scalar 3; Scalar 4]; & \\
+        & Matrix [Scalar 5; Scalar 6]] |\} &
+    \end{tabular}
+    \caption{Example matrices and their Coq equivalents}\label{T:matrices}
+\end{table*}
+
+\subsubsection{The well-formed property}
+
+As suggested above, the lack of dependent typing means a value of type
+\(\func{matrix}\, \var{A}\) is \emph{not} guaranteed to be well-formed. All of
+the following are valid but nonsensical values:
+
+\begin{verbatim}
+{| shape := [];
+   contents := Matrix [Scalar 1] |}
+{| shape := [2];
+   contents = Matrix [Scalar 2] |}
+{| shape := [1; 2; 3; 0];
+   contents = Scalar 0 |}
+\end{verbatim}
+
+In the first case, we have a matrix with no dimensions, which should be a scalar
+and is not. In the second case, we have a one-dimensional matrix of two
+elements which only contains one element. In the third, we have an empty
+four-dimensional matrix which is actually a scalar. None of these pairings makes
+sense---we need a way to exclude them from our results, effectively categorizing
+these matrices as ``not well-formed.'' Conversely, we would like a way to
+categorize matrices as well-formed.
+
+We have informally given pieces of the definition in laying our the relationship
+between shape and contents; here, we make the definition more precise. A matrix
+is well-formed if, and only if,
+\begin{itemize}
+    \item the shape is an empty list, and the contents are a single scalar value;
+        or
+    \item the shape is a non-empty list consisting of a head \(n\) representing
+        the length of the first dimension and tail \(t\) representing the shape
+        of the subsequent dimensions, and the contents is a \(\func{Matrix}\)
+        containing a list with length \(n\), and each element of the list is
+        well-formed with respect to \(t\).
+\end{itemize}
+In Coq, we express this definition via a recursive definition
+\(\func{well\_formed'}\) which computes a proposition, or proof-burden, over a
+shape and contents. We wrap this up in a definition \(\func{well\_formed}\),
+which unpacks a matrix for use with \(\func{well\_formed'}\). With this
+definition, we can theorems in terms of matrices that are assumed to be
+well-formed or that must be shown to be well-formed.
+
+In the course of our development, we discovered that this computational version
+did not always provide a convenient means to work a proof. Clearly, the
+recursive nature of matrices and well-formedness means that most proofs will be
+by induction. But induction on what? There is no induction principle for the
+record type, so we must choose from induction on the shape or on the contents.
+It turns out that there is a natural inductive version of the well-formed
+property that permits induction on evidence. In \S~\ref{S:matrix_thm} we will
+state and sketch the equivalence theorem between the two definitions. For now,
+we give the judgment rules for \(\func{well\_formedI'}\), the inductive
+counterpart to \(\func{well\_formed'}\) in \figurename~\ref{F:wfI}. We wrap the
+inductive proposition in \(\func{well\_formedI}\), which unpacks a matrix the
+same way as \(\func{well\_formed}\).
+
+\begin{figure*}[ht]
+    \centering
+    \begin{mathpar}
+        \inference[\iname{WF\_Scalar}]{%
+            a: A
+        }{\func{well\_formedI'}\, []\, (\func{Scalar}\, a)}
+
+        \and
+
+        \inference[\iname{WF\_Empty}]{%
+            t: \func{list}\, \mathbb{N}
+        }{\func{well\_formedI'}\, (0::t)\, (\func{Matrix}\, [])}
+
+        \and
+
+        \inference[\iname{WF\_Cons}]{%
+            h: \mathbb{N} &
+            t: \func{list}\, \mathbb{N} &
+            m: \func{matrix\_contents}\, \var{A} \\
+            ms: \func{list}\, (\func{matrix\_content}\, \var{A}) &
+            \func{well\_formedI'}\, (h::t)\, (\func{Matrix}\, ms) &
+            \func{well\_formedI'}\, t\, m
+        }{\func{well\_formedI'}\, (1+h::t)\, (\func{Matrix}\, (m::ms))}
+
+    \end{mathpar}
+    \caption{Inductive judgments for the well-formed property of
+    matrices}\label{F:wfI}
+\end{figure*}
+
+The important features of the judgements are \(\iname{WF\_Emtpy}\) and
+\(\iname{WF\_Cons}\). The judgement \(\iname{WF\_Empty}\) corresponds to the
+vacuous case: all elements in an empty list are well-formed with respect to any
+shape (there are no such elements!). The judgment \(\iname{WF\_Cons}\) provides
+a judgment for increasing the size of the first dimension of a well-formed
+matrix by adding a well-formed component to the contents.
+
+It is easy to prove that all the matrices in Table~\ref{T:matrices} are
+well-formed with respect to both definitions. It is likewise easy to prove that
+the previously-sketched invalid examples are not well-formed under both
+definitions.
+
+\subsubsection{Some operations on matrices}
+
+The operations we have defined are computing the shape of a matrix, linearizing
+the matrix, and indexing a multi-dimensional matrix linearly. We will discuss
+these and give a brief note about our approach to multi-dimensional range-based
+indexing.
+
+\paragraph{Computing the shape of a matrix} While each matrix carries its shape
+as part of the record pairing, it is occasionally convenient to compute from the
+contents alone the shape. There are two subtleties worth mentioning: first, the
+computation is only valid for well-formed matrices because it relies on
+assumptions about the relationship between different components of the contents.
+Second, the computed shape may not be the most general: we have already seen
+that, when one dimension in a matrix has shape zero, the subsequent dimensions
+can be arbitrarily sized and the result is still well-formed! This impacts the
+correctness statement for shape-computation: on a well-formed matrix, it may not
+give the exact same shape, yet it will give a shape such that the original
+matrix is well-formed with respect to the new shape. We can call this
+``normalization'' of the shape, as it discards dimensions that follow a
+zero-sized dimension. See \S~\ref{S:matrix_thm} for more details on
+normalization via shape computation. The definition is straightforward: the
+function \(\func{compute\_shape}\) yields
+\begin{itemize}
+    \item an empty list when the contents consist of a single scalar;
+    \item a singleton list containing zero when the contents consist of an empty
+        matrix; and
+    \item a list with head the length of the matrix contents and tail the
+        computed shape of just one of the contents when the contents
+        consist of a non-empty matrix.
+\end{itemize}
+This last bullet explains why the function is only valid on well-formed
+matrices: it implicitly assumes that all the elements of the matrix contents
+have the same computed shape, which is only true when the matrix is well-formed.
+The same definition is given below:
+
+\begin{align*}
+    & \func{compute\_shape}(\var{m}) \triangleq \\
+    & \begin{cases}
+        [] & m = \func{Scalar}\, a \\
+        [0] & m = \func{Matrix}\, [] \\
+        |\var{ms}| :: \func{compute\_shape}\, m' & m = \func{Matrix}\, \var{ms}, \\
+        & \var{ms} = m'::t
+    \end{cases}
+\end{align*}
+
+\paragraph{Linearizing a matrix} See \S~\ref{S:design_linearize} for more
+details on matrix linearization. To define linearization in Coq, we use a
+two-stage unpacking as with other matrix operations. The function
+\(\func{linearize'}\) turns scalars into a singleton list and recursively
+flat-maps itself over matrices to produce the one-dimensional matrix as a list.
+The function \(\func{linearize}\) unpacks a matrix, then transforms the contents
+by wrapping the result of mapping \(\func{linearize'}\) into \(\func{Scalar}\)s
+in a \(\func{Matrix}\) and transforms the shape by taking the product of the
+(normalized) shape of the matrix. Recall that the length of the linearized
+vector is the product of the sizes of all the dimensions. The definitions are
+spelled out below:
+
+\begin{align*}
+    & \func{linearize'}(\var{contents}) \triangleq \\
+    & \begin{cases}
+        [\var{a}] & \var{contents} = \func{Scalar}\, \var{a} \\
+        \func{flat\_map}\, \func{linearize'}\, \var{ms} & \var{contents} =
+        \func{Matrix}\, ms
+    \end{cases} \\
+    & \func{linearize}(\{\var{contents}, \var{shape}\}) \triangleq \\
+    & \{ \var{shape} = \left[\prod \func{compute\_shape}\, \var{contents}\right]; \\
+    & \ \, \var{contents} = \func{Matrix}\, (\func{map}\, \func{Scalar}\, (\func{linearize'}\, \var{contents})) \}
+\end{align*}
+
+\paragraph{Linearly indexing a matrix} Given the operation to linearize a
+matrix, one-based linearly indexing is easy. See \S~\ref{S:design_linear_index}
+for details on linear indexing. The function \(\func{nth}\) takes a matrix \(m\)
+and an index \(n\); it returns the one-based linearly indexed content of \(m\)
+at \(n\) (wrapped in the option constructor \(\func{Some}\)) if \(n\) is in
+bounds relative to \(m\). When \(n\) is out of bounds, the result is
+\(\func{None}\). For one-based indexing with functions that naturally use
+zero-based indexing, we use \(n-1\). We should note that the current version of
+\(\func{nth}\) has the subtle bug that indexing at \(n=0\) currently returns the
+same as indexing at \(n=1\), when it should in fact be out of bounds and thus
+\(\func{None}\)---a simple \texttt{if} or \texttt{match} guard would cover this
+case.
+
+\subsection{Key Theorems about Matrices}\label{S:matrix_thm}
+
+\subsection{Advanced Coq for Matrices}\label{S:matrix_coq}
+
 {\printbibliography}

### 2021-05-02 paper/paper.tex
paper: start theorems

@@
 \usepackage{array,booktabs}
+\usepackage{amsthm}
 \usepackage{graphicx}
@@
 \newcommand{\iname}[1]{\textsf{#1}}
+
+\theoremstyle{plain} % default
+\newtheorem{thm}{Theorem}[section]
+\newtheorem{lem}[thm]{Lemma}
+\newtheorem{prop}[thm]{Property}
+\newtheorem*{cor}{Corollary}
+
+\theoremstyle{definition}
+\newtheorem{defn}{Definition}[section]
+\newtheorem{example}{Example}[section]
+\newtheorem{exercise}{Exercise}[section]
+\newtheorem*{prob}{Problem}
+
+\theoremstyle{remark}
+\newtheorem*{rem}{Remark}
+\newtheorem*{note}{Note}
+\newtheorem{case}{Case}
 
@@
 
-The important features of the judgements are \(\iname{WF\_Emtpy}\) and
+The important features of the judgements are \(\iname{WF\_Empty}\) and
 \(\iname{WF\_Cons}\). The judgement \(\iname{WF\_Empty}\) corresponds to the
@@
 
-\paragraph{Linearly indexing a matrix} Given the operation to linearize a
+\paragraph{Linearly indexing a matrix} Given the operation to linearise a
 matrix, one-based linearly indexing is easy. See \S~\ref{S:design_linear_index}
@@
 
+Unless otherwise stated, the key theorems are predicated on all matrices
+involved being well-formed.
+
+The key theorems of our proof development are
+\begin{inlist}
+\item that the two different definitions of the well-formed property agree
+    (Theorem~\ref{Th:well_formed_agree});
+\item that \(\func{compute\_shape}\) normalizes by preserving the well-formed
+    property (Theorem~\ref{Th:compute_shape_wf_normalizes}); and
+\item that \(\func{linearize}\) is correct, \textit{i.e.}, that
+    \(\func{linearize'}\) produces a list with size the product of the shape and
+    that \(\func{linearize}\) preserves the well-formed property
+    (Theorems~\ref{Th:linearize_product} and~\ref{Th:linearize_wf}).
+\end{inlist}
+We present these, along with a sketch of their proofs, in order.
+
+\begin{thm}[Well-formed definitions agree]\label{Th:well_formed_agree}
+    For all matrices \(m\), \(m\) is well-formed under the computational
+    definition \(\func{well\_formed}\) if-and-only-if \(m\) is well-formed under
+    the inductive definition \(\func{well\_formedI}\). (See
+    \texttt{well\_formed\_agree} in the Coq
+    development~\cite{zelda_mosaic_proof}.)
+\end{thm}
+
+\begin{proof}
+    We begin with the forward direction and proceed by induction on the contents
+    of \(m\); for each case of the contents, we also consider the cases of the
+    shape. Several cases are immediately contradictory, since the resulting
+    \(m\) could not have been well-formed in the first place. In the remainder,
+    we must prove \(\func{well\_formed}\, m\) implies \(\func{well\_formedI}\,
+    m\).
+
+    The base case is easy: simply use \(\iname{WF\_Empty}\).
+
+    In the inductive case, we can use the well-formed hypothesis together with
+    the induction hypothesis and auxiliary Lemma~\ref{Lem:wfI_all_wf_t} to
+    finish.
+
+    Now the backwards direction. We must prove that \(\func{well\_formedI}\, m\)
+    implies \(\func{well\_formed}\, m\). We proceed by induction on the evidence
+    for the hypothesis \(\func{well\_formedI}\, m\).
+
+    The base cases are straightforward and follow readily from the definition.
+
+    In the inductive case, we can use the inductive hypotheses to conclude.
+\end{proof}
+
+\begin{lem}\label{Lem:wfI_all_wf_t}
+    If all elements of the list \(\var{ms}\) are well-formed (using the inductive
+    definition) with respect to some shape \(t\), then the following matrix is
+    also well-formed: \(\{ \var{shape} = |\var{ms}| :: t; \var{contents} =
+    \func{Matrix}\, \var{ms} \}\).
+\end{lem}
+
+The proof is straightforward by induction on \(\var{ms}\).
+
+\begin{thm}[Computing the shape normalizes it]\label{Th:compute_shape_wf_normalizes}
+    For all matrices \(m\), if \(m\) is well-formed than so is the following
+    matrix: \(\{ \var{shape} = \func{compute\_shape}\, m; \var{contents} =
+    \func{contents}\, m \}\).
+\end{thm}
+
+\begin{proof}
+    We proceed by induction on the contents of \(m\); for each case of the
+    contents, we also consider the cases of the shape. Several cases are
+    immediately contradictory, since the resulting \(m\) could not have been
+    well-formed in the first place.
+
+    The base case is straightforward from the definition of
+    \(\func{well\_formed}\).
+
+    In the inductive case, we have to deal with a \(\func{Matrix}\) of some list
+    \(\var{ms}\). We proceed by case analysis on the list; the empty case is
+    easy. When the list is non-empty, we are left to prove the critical
+    assumption in the definition of \(\func{compute\_shape}\); namely, that all
+    the elements in a well-formed matrix have the same computed shape. This is
+    done by application of the induction hypothesis and observing that any
+    element in the non-empty \(\var{ms}\) has the same computed shape, thanks to
+    Lemma~\ref{Lem:wf_same_shape}.
+\end{proof}
+
+\begin{lem}\label{Lem:wf_same_shape}
+    Given any two matrices \(m_1, m_2\) that are well-formed with respect to the
+    same shape, \(\func{compute\_shape}\, m_1 = \func{compute\_shape}\, m_2\).
+\end{lem}
+
+The proof is straightforward by induction on the shape combined with case
+analysis on the two matrices.
+
+\begin{thm}[Linearize has resulting shape the product of the original]\label{Th:linearize_product}
+\end{thm}
+
+\begin{thm}[Linearize preserves the well-formed property]\label{Th:linearize_wf}
+\end{thm}
+
 \subsection{Advanced Coq for Matrices}\label{S:matrix_coq}

### 2021-05-02 paper/paper.tex
paper: copy overleaf diffs over

@@
 \usepackage[inline]{enumitem}
+\usepackage{tikz}
+\usetikzlibrary{matrix,calc}
 \newlist{inlist}{enumerate*}{1}
@@
 
-\section{Progress Report}
+\section{Problem Description}
+Many non-trivial programs make use of arrays, and our \texttt{zelda-mosaic} program is no different.
+Because our goal was to prove the termination of the program when given a well-formed input and the core data structure modified throughout the program was an image (represented as an array), it is helpful to first describe the array operations used and the invariants of an array that contribute to ``well-formed-ness".
 
-The starting point to our project was to fill in some knowledge gaps that we had
-not covered in class prior to when we began. Since we planned to model a
-program, we needed to know more about the modeling of programs and execution. To
-this end, we completed work in \textit{Software Foundations} up to the chapter
-\textit{Simple Imperative Programs}~\cite{Pierce:SF1}. This chapter provided
-great insight on how to model statements and expressions, maintain state over a
-program's execution, and define semantics. After completing this, we were better
-able to outline the scope of the project.
+\subsection{Array Usage}
+\matlab's support for multi-dimensional arrays makes it amenable to image-processing tasks such as ours.
+The algorithms we wrote are described at high-level in pseudocode in Figures \figurename~\ref{alg:v1}, \ref{alg:v2}, and \ref{alg:v3}, but here we explain them in layman's terms to help motivate some features of our proof.\\
 
-Our progress thus far can be divided into two main categories:
-\begin{inlist}
-\item determining the subset of \matlab\ (dubbed \mmatlab) relevant to our program, and
-\item creating data structures and functions suitable to model the behaviors of
-    \mmatlab\@.
-\end{inlist}
+As described in our pseudocode, the inputs to our program were a piece of \textit{key art} to be recomposed with \textit{thumbnails} of a certain \textit{size}, defined as the width and height of the thumbnails (all have been coerced to be square-shaped).
+The original representation of \textit{key art} in \matlab is a two-dimensional array (with indices corresponding to a particular pixel's row and column) containing an additional array of three values - one each for the red, green, and blue color channels.
+So far, this is an array of three dimensions.
+We then divide this image into \textit{chunks}, where each \textit{chunk} will be replaced by a \textit{thumbnail} in the final product.
+For \textit{key art} of dimensions $m \times n$ (with $m$ corresponding to and $n$ corresponding to columns), each image will be divided into $(m / size) \times (n / size)$ \textit{chunks}, which will each be replaced by a \textit{thumbnail} as appropriate.
+These \textit{chunks}, each of size $size \times size$, are then stored in a two-dimensional array, with each \textit{chunk} retaining its original placement in the \textit{key art}.
+As a result we now have a five-dimensional array, and in order to access a particular value from a color channel, we need the following values:
 
-We also identify two main future tasks that will rely upon the completion of
-these tasks:
-\begin{inlist}
-\item binding operations from the \mmatlab\ language to semantics in
-    Coq~\cite{Coq}, and
-\item using Hoare logic~\cite{Hoare_1969} to prove that the program terminates
-    under what pre-conditions.
-\end{inlist}
+\begin{itemize}
+\item the \textit{chunk row} it is in
+\item the \textit{chunk column} it is in
+\item the \textit{x-offset} of the pixel into its chunk
+\item the \textit{y-offset} of the pixel into its chunk
+\item the \textit{color channel} desired out of red, green, and blue
+\end{itemize}
+
+Thus, once a piece of \textit{key art} is successfully divided into \textit{chunks}, the image is effectively represented as an array whose dimensions correspond to the above values.
+The \textit{chunks} are iterated through in row-major order, gradually being replaced in-place by \textit{thumbnails}.
+Consequently, the state of our core array maintains some matrix-like properties that are supported by \matlab\@ arrays, but not particular to \matlab\@ in general:
+
+\begin{itemize}
+    \item The shape of a matrix \textbf{$M$} can be described by a list of \textit{d} dimensions, where each dimension's value $v_i$ refers to the number of elements held along that dimension.
+For example, an array with dimensions $3 \times 4 \times 2$ contains 3 arrays, each of which contains 4 arrays of 2 elements.
+$v_1 = 3$, $v_2 = 4$, and $v_3 = 2$.
+    \item Once the shape of \textbf{$M$} is defined and all of the elements have been populated, the shape will not change and all of the elements will remain populated.
+For example, the aforementioned $3 \times 4 \times 2$ array must always contain 3 arrays, each of which contains 4 arrays of 2 elements.
+    \item Each $i$-th dimension from $1$ to $d-1$ must contain $v_i$ elements, each of which is an array of length $v_{i+1}$; in other words, the length of each element in a particular dimension must be identical to one another.
+\end{itemize}
+
+In summary, these properties enforce a notion of ``well-formed-ness" that ensures that the size $v_i$ of each dimension remains fixed and that all element in a particular dimension $i$ have the same length as one another.
+This property is generally true of matrices, hence the references to both an ``arrays" and ``matrices" in this section.
+Unless otherwise noted, ``array" refers to the generic \matlab\@ object and ``matrix" refers to an ``array" with these properties when discussed in the scope of this paper.
+
+\subsection{Operations} \matlab\@ supports a number of operations that can be used to manipulate arrays, but only a subset are used in our program.
+To illustrate the operations used, we begin with the following visualization of a three-dimensional array \textbf{$E$} with dimensions $3 \times 4 \times 2$, as it is used in \matlab:\\
+
+% Code adapted from: https://tex.stackexchange.com/questions/300109/simple-visualization-of-3d-matrix
+
+\begin{figure}[h!]
+    \centering
+    \begin{tikzpicture}[every node/.style={anchor=north east,fill=white, minimum width=.7cm}]
+    \matrix (mA) [draw,matrix of math nodes]
+    {
+    1 & 3 & 5 & 7 \\
+    2 & 4 & 6 & 8 \\
+    };
+
+    \matrix (mB) [draw,matrix of math nodes] at ($(mA.south west)-(-1.5, .1)$)
+    {
+    9 & 11 & 13 & 15 \\
+    10 & 12 & 14 & 16 \\
+    };
+
+    \matrix (mC) [draw,matrix of math nodes] at ($(mB.south west)-(-1.5, .1)$)
+    {
+    17 & 19 & 21 & 23 \\
+    18 & 20 & 22 & 24 \\
+    };
+
+    \draw[dashed](mA.north east)--(mC.north east);
+    \draw[dashed](mA.north west)--(mC.north west);
+    \draw[dashed](mA.south east)--(mC.south east);
+    \end{tikzpicture}
+
+\end{figure}
+
+\subsubsection{Element Access}
+For the matrix \textbf{$E$}, each element would be indexed according to the following mapping.
+Under this scheme, for example, the element $8$ in \textbf{$E$} would have the index $(1, 4, 2)$.
+This indexing scheme is the general pattern used to access one particular element out of a multi-dimensional array.
+Note that arrays in \matlab\@ have indices starting at 1.
+
+\begin{figure}[h!]
+    \centering
+    \begin{tikzpicture}
+    \def\xs{1} %shift in x direction
+    \def\ys{1.5} %shift in y direction
+    \def\nm{3} % number of 2d matrices in the 3d matrix
+    \foreach \x in {1,2,...,\nm}
+    {
+    \matrix [draw, % for the rectangle border
+             fill=white, % so that it is not transparent
+             ampersand replacement=\&] %see explanation
+    (mm\x)%give the matrix a name
+    at(-\x * \xs, -\x * \ys) %shift the matrix
+    {
+        \node {(\x,1,1)}; \& \node {(\x,2,1)}; \& \node {(\x,3,1)}; \& \node {(\x,4,1)};\\
+        \node {(\x,1,2)}; \& \node {(\x,2,2)}; \& \node {(\x,3,2)}; \& \node {(\x,4,2)};\\
+    };
+    }
+
+    \draw [dashed,gray](mm1.north west) -- (mm\nm.north west);
+    \draw [dashed,gray](mm1.north east) -- (mm\nm.north east);
+    \draw [dashed,gray](mm1.south east) -- (mm\nm.south east);
+    \end{tikzpicture}
+\end{figure}
+
+\subsubsection{Linearized Indexing} \label{linearizedindexing}
+Note that in $E$, elements are ordered in ascending order with respect to their indexing.
+First, sub-array $(1)$ is populated, and $(1, 1)$ is the first array within it, so it is filled with elements $1$ and $2$.
+Array $(1, 2)$ is then populated by elements $3$ and $4$.
+After all four of \textbf{$E$}'s sub-arrays have been populated, sub-array $(2)$ is populated, and so on until the array is full.
+
+Given the numbering scheme in $E$, one can imagine a ``flattened" version of $E$ that is simply an ordered one-dimensional array containing all of $E$'s elements; simply put, an array containing the elements $1$ to $24$, in numerical order.
+\matlab\@ actually allows arrays of arbitrarily many dimensions to be logically ``flattened" during accesses so that elements can be accessed with one index, like in a one-dimensional array.
+We found this property to be somewhat interesting and challenging to model, and we refer to the ``flattening" of a matrix as ``linearization".
+
+The process of linearizing a matrix is perhaps
+better explained with a two-dimensional example. Imagine a general two-dimensional $3 \times 4$ matrix, such as the one below, where:
+\begin{equation*}
+    \begin{bmatrix}
+        a & b & c & d \\
+        e & f & g & h \\
+        i & j & k & l
+    \end{bmatrix}
+\end{equation*}
+To linearize the matrix, we label each value with an increasing index from the
+top-left, left-to-right, top-to-bottom. For example:
+\begin{equation*}
+    \begin{bmatrix}
+        a_1 & b_2 & c_3 & d_4 \\
+        e_5 & f_6 & g_7 & h_8 \\
+        i_9 & j_{10} & k_{11} & l_{12}
+    \end{bmatrix}
+\end{equation*}
+Finally, we create a vector, a one-dimensional matrix, by reading the elements
+off in order:
+\setcounter{MaxMatrixCols}{20}
+\begin{equation*}
+    \begin{bmatrix}
+        a & b & c & d & e & f & g & h & i & j & k & l
+    \end{bmatrix}
+\end{equation*}
+This process generalizes to higher dimensions. The length of the resulting
+vector is always the product of the sizes of the dimensions. To see this, note
+that this product counts the number of elements in the matrix and that the
+linearized matrix has the same number of elements in a single dimension.
+
+Under this scheme, any element in a matrix can be accessed via its position in the linearized version of that matrix.
+For example, the element $12$ in \textbf{$E$} can be accessed by the index $(12)$ (note that 12 refers to the index, not the elements) and in the previous example, the element $f$ can be accessed by the index $(6)$.
+
+\subsubsection{Ranged Indexing} \label{rangedindexing}
+Broadly speaking, \matlab\@ supports the notion of a range of numbers specified by a starting number and a stopping number, which includes every number between the starting number and the stopping number (inclusive).
+This abstraction can be used to access or assign elements within a sub-matrix of a particular matrix.
+For example, suppose in matrix $E$ you only cared about the first matrix in the first dimension, the second/third/fourth columns within the second dimension, and the second row within the third dimension.
+This sub-matrix can be specified by the index $(1, 2:4, 2)$ and would return the matrix:
+\begin{equation*}
+    \begin{bmatrix}
+        4 & 6 & 8
+    \end{bmatrix}
+\end{equation*}
+
+Suppose that you wanted to get the second/third/fourth columns within the second dimension and second row within the third dimension for \textit{all three} of the outer-most dimension of \textbf{$E$} but did not know a priori how many elements it had.
+\matlab\@ supports an ``inclusive range" operator, represented by a colon \textbf{:}, which is taken to mean the range encompassing all values in a particular dimension.
+The sub-matrix described here would be specified by the index $(:, 2:4, 2)$ and represents the following matrix:
+
+\begin{figure}[h!]
+    \centering
+    \begin{tikzpicture}[every node/.style={anchor=north east,fill=white, minimum width=.7cm}]
+    \matrix (mA) [draw,matrix of math nodes]
+    {
+    4 & 6 & 8 \\
+    };
+
+    \matrix (mB) [draw,matrix of math nodes] at ($(mA.south west)-(-1.5, .1)$)
+    {
+    12 & 14 & 16 \\
+    };
+
+    \matrix (mC) [draw,matrix of math nodes] at ($(mB.south west)-(-1.5, .1)$)
+    {
+    20 & 22 & 24 \\
+    };
+
+    \draw[dashed](mA.north east)--(mC.north east);
+    \draw[dashed](mA.north west)--(mC.north west);
+    \draw[dashed](mA.south east)--(mC.south east);
+    \end{tikzpicture}
+
+\end{figure}
+
+\subsubsection{Matrix Operations} \label{matrixoperations}
+Details about \matlab\@ arrays pertaining to indexing are perhaps some of their most interesting features, but we also use some typical matrix operators in our program.
+Generally speaking, matrix operators can be generalized to any number of dimensions, as long as the dimensions of the arrays agree for the operation in question.
+There are two broad classes of matrix operators that we make use of:
+
+\begin{itemize}
+    \item \textbf{Element-wise}: For two matrices of the same dimensionality, the operator is applied on each of the two elements containing the same index in the two matrices.
+    \item \textbf{All Elements}: For an array and a scalar, the operation and scalar (e.g. add 1) is applied to every element in the matrix.
+\end{itemize}
+
+\subsection{Modeling the Problem}
+Aside from the core matrix operations, we anticipated that we would need to model the state of our program at all points of execution since our goal was to prove termination.
+To this end, we would need to identify all of the syntax that we use in our program and specify their associated semantics.
+The upshot of modeling our program in this way would be the ability to remains to prove its termination of our given that certain preconditions hold.
+
+Hoare logic~\cite{Hoare_1969} is a type of logic that is often used to solve problems of this class.
+Partial Hoare logic is a weak statement about programs; namely, given preconditions \(P\) and postconditions \(Q\) and a program \(C\), the Hoare triple \(\{P\}C\{Q\}\) asserts that \emph{if} \(C\) terminates, then the pre-conditions imply the post-conditions.
+Total Hoare logic, by contrast, is a stronger version of this claim: \(C\) is asserted to terminate.
+Thus a proof in total Hoare logic necessitates a proof that \(C\) terminates; the inference rules are accordingly adjusted.
+The primary adjustment is to the rule for while
+loops, which are guaranteed to terminate by relation to a decreasing sequence
+(with respect to a well-founded order).
+
+When we started this project, we had in mind that the preconditions would be statements about the program's inputs; for example, ``the length and width of \textit{key art} are always multiples of the \textit{thumbnails'} size".
+Ultimately, we did not get far enough in modeling our program to devise a set of pre-conditions, but the point remains that we needed to be able to model the control flow of our entire program in order to do this.
+
+\section{Design / Approach}
+
+We first set out to define the syntax and semantics of the subset of \matlab\ used in our program, making the assumption that once we finished this it would be somewhat trivial to bind the defined syntax to behavior in Coq.
+Ultimately, the majority of our effort in this project ended going towards the implementation of our multi-dimensional matrix operations, which was made difficult by the number of properties that our matrices uphold as well as our desire to write generic Coq functions that would work on matrices of arbitrarily many dimensions.
+We spent a while working with Coq's built-in \textsf{Vector} type before coming to the conclusion that its dependent typing caused difficulties in behavior specification.
+To this end, we ended up implementing our own data types to describe the state of matrices in our program; this task was unexpectedly challenging and once we began working on it, it became the only thing we worked on for the remainder of the project.
 
@@
 
-We give a brief overview on \mmatlab, the subset of \matlab\ that we model in
+We present a brief overview on \mmatlab, the subset of \matlab\ that we modeled in
 order to prove facts about our programs.
@@
 operations, and \matlab-specific primitives such as ranges. We have identified
-all of the expressions and statements we expect to use and defined them as
+all of the expressions and statements we expected to use and defined them as
 inductive types, shown in \figurename~\ref{fig:inductivetypeexp} and
@@
         matrix, is a matrix with all elements incremented by the number. Similar
-        behaviors hold for \textsf{MultExp} and \textsf{DivExp}.
+        behaviors hold for \textsf{MultExp} and \textsf{DivExp}. See Section \ref{matrixoperations} for the distinction between these operators.
 
@@
         starting number and stopping number may be provided, generating an
-        inclusive list of all numbers between them.
+        inclusive list of all numbers between them. This is discussed in Section \ref{rangedindexing}
+
+    \item \textsf{IndexExp}: In support of the RangeExp expression, the syntax to access elements from an array is overloaded in \matlab\@. They
 
@@
 \end{itemize}
-
-\subsection{Data Structures and Functions}
-
-Our program is used for image processing; thus, many computations are made on
-images, which are represented as arrays in \matlab\@. Additionally, we deal with
-matrices of images, so we need to model multi-dimensional matrices. Since
-multi-dimensional matrices can have arbitrarily many dimensions, we decided the
-best abstraction for this in Coq would be built on top of the \textsf{Vector}
-type which was originally implemented in the \colorlib\
-library~\cite{BLANQUI_2011} and eventually ported to the Coq standard library.
-In Coq, \textsf{Vector} is a dependent type~\cite{Bove2009,Thorsten_2010}; the
-number of elements in a \textsf{Vector} is part of its type, written
-\texttt{Vector.t A n} for a vector containing \(n\) elements of type \(A\).
-Since \textsf{Vector}s are polymorphic, we noted that a \textsf{Vector} could
-contain \textsf{Vector}s provided they are all the same type (which includes
-length, as noted). That is, we can stack \textsf{Vector}s to form, \emph{e.g.},
-\texttt{Vector.t (Vector.t A m) n}. From there, we defined a type-constructor
-\textsf{Matrix} taking a list of dimensions, which represents a \textsf{Vector}
-of \textsf{Vector}s. It can be used as the type of an arbitrary
-\(n\)-dimensional matrix containing elements of type \(A\).
-
-Due to the implementation of \textsf{Vector} in the Coq standard library, we
-needed to greatly extend the functionality of our \textsf{Matrix} type. To that
-end, we have begun implementing \textsf{Matrix} functions such as element
-access, which is non-trivial for matrices with an arbitrary number of
-dimensions\footnote{In particular because of the dependent types; see also
-\url{https://stackoverflow.com/q/66863226/4400820}}. Additionally, the existing
-implementation of element access in Coq's \textsf{Vector} type uses the
-unfamiliar \textsf{Fin} type, an abstraction for carrying proofs that a
-particular number belongs to finite range of numbers. In order to access an
-element of a \textsf{Vector}, one must supply such a proof; for our purposes, we
-use \texttt{Fin.of\_nat}, which gives either such a proof or a witness that the
-number is outside the range. This behavior is captured by the \textsf{sumbool}
-built-in, which required further research on our behalf. Given the number of
-ways that matrices can be indexed in \matlab\ and the complexity introduced by
-Coq's implementation of \textsf{Vector}s, we see this as the main bottleneck for
-our progress. We still need to implement many \textsf{Matrix} operations, such
-as accesses over ranges and over an entire dimension.
-
-\subsection{Binding Operations and Semantics}
-
-While we have not yet bound any of our functions or data types to the statements
-or expressions in \S~\ref{S:mmatlab}, we have naturally been thinking about them
-throughout the whole process. The functions we write are intended to implement
-the evaluation of the corresponding expressions. In the style of
-\texttt{Imp}~\cite{Pierce:SF1}, we intend to write a deterministic expression
-evaluation function; we will then provide semantics for program execution.
-Execution cannot be defined as a function in Coq because it may be
-non-terminating. We have had to consider what operations are be allowed and
-whether checks for undesired behavior (\emph{e.g.}, dynamic type mismatches)
-should be implemented in the underlying functions or in a semantic check.
-
-\subsection{Hoare Logic}
-
-We have a significant amount of work to do in modeling \mmatlab, but the
-blue-sky goal remains to prove the termination of our program given that certain
-preconditions hold. Ben has been studying Hoare logic~\cite{Hoare_1969}, the
-relevant background material for this problem, as part of his Thesis
-requirement. What will be possible to prove will be determined in part by what
-we are able to successfully model, given the time constraints. However, as a
-preliminary step, we have studied partial Hoare logic and total Hoare logic to
-provide more context for the task ahead of us. In this exploration, we found
-that partial Hoare logic is a weak statement about programs; namely, given
-preconditions \(P\) and postconditions \(Q\) and a program \(C\), the Hoare
-triple \(\{P\}C\{Q\}\) asserts that \emph{if} \(C\) terminates, then the
-pre-conditions imply the post-conditions. Total Hoare logic, by contrast, is a
-stronger version of this claim: \(C\) is asserted to terminate. Thus a proof in
-total Hoare logic necessitates a proof that \(C\) terminates; the inference
-rules are accordingly adjusted. The primary adjustment is to the rule for while
-loops, which are guaranteed to terminate by relation to a decreasing sequence
-(with respect to a well-founded order). Given that we are trying to prove
-termination, it seems likely that we will need to use total Hoare logic in the
-relevant proofs.
-
-\begin{figure}[ht]
-    \textbf{Input}: image \(key\_art\), images \(thumbnails\), integer \(size\) \\
-    \textbf{Output}: image \(tiles\)
-    \begin{algorithmic}
-        \STATE{\(tiles \gets \textrm{MAT2TILES}(key\_art, size, size)\)}
-        \FOR{\(tile \in tiles\)}
-            \FOR{\(thumbnail, i \in thumbnails, [1..|thumbnails|]\)}
-                \STATE{\(mses_i \gets \textrm{immse}(tile, thumbnail)\)}
-            \ENDFOR
-            \STATE{\(best\_indices \gets \textrm{find} (mses = \min mses)\)}
-            \STATE{\(best\_index \gets best\_indices_0\)}
-            \STATE{\(tile \gets thumbnails_{best\_index}\)}
-        \ENDFOR
-    \RETURN{\(tiles\)}
-    \end{algorithmic}
-    \caption{Algorithm: Version 1}\label{alg:v1}
-\end{figure}
-
-\begin{figure}[ht]
-    \textbf{Input}: image \(key\_art\), images \(thumbnails\), integer \(size\) \\
-    \textbf{Output}: image \(tiles\)
-    \begin{algorithmic}
-        \STATE{\(tiles \gets \textrm{MAT2TILES}(key\_art, size, size)\)}
-        \STATE{\(used \gets \emptyset\)}
-        \STATE{\(threshold \gets 1\)}
-        \IF{\(|thumbnails| < |tiles|\)}
-            \STATE{\(threshold \gets \lceil{|tiles| / |thumbnails|}\rceil\)}
-        \ENDIF
-        \FOR{\(tile \in tiles\)}
-            \FOR{\(thumbnail, i \in thumbnails, [1..|thumbnails|]\)}
-                \STATE{\(mses_i \gets \textrm{immse}(tile, thumbnail)\)}
-            \ENDFOR
-            \STATE{\(best\_indices \gets \textrm{find} (mses = \min mses)\)}
-            \STATE{\(best\_index \gets best\_indices_0\)}
-            \WHILE{\(used_{best\_index} \ge threshold\)}
-                \STATE{\(mses_{best\_index} \gets \infty\)}
-                \STATE{\(best\_indices \gets \textrm{find} (mses = \min mses)\)}
-                \STATE{\(best\_index \gets best\_indices_0\)}
-            \ENDWHILE
-            \STATE{increment \(used_{best\_index}\)}
-            \STATE{\(tile \gets thumbnails_{best\_index}\)}
-        \ENDFOR
-        \RETURN{\(tiles\)}
-    \end{algorithmic}
-    \caption{Algorithm: Version 2}\label{alg:v2}
-\end{figure}
-
-\begin{figure}[ht]
-    \textbf{Input}: image \(key\_art\), images \(thumbnails\), integer \(size\) \\
-    \textbf{Output}: image \(tiles\)
-    \begin{algorithmic}
-        \STATE{}\COMMENT{Repeat Version 2}
-        \STATE{\(tiles \gets \textrm{Version2}(key\_art, thumbnails, size)\)}
-        \STATE{\(tiles \gets \textrm{imgaussfilt}(tiles)\)}
-        \STATE{\(strip\_size \gets \lfloor{size / 8}\rfloor\)}
-        \STATE{}\COMMENT{Blur by averaging}
-        \FOR{horizontal \(strip \in tiles\)}
-            \STATE{\(strip \gets \textrm{mean}({strip})\)}
-        \ENDFOR
-        \FOR{vertical \(strip \in tiles\)}
-            \STATE{\(strip \gets \textrm{mean}({strip})\)}
-        \ENDFOR
-        \RETURN{\(tiles\)}
-    \end{algorithmic}
-    \caption{Algorithm: Version 3}\label{alg:v3}
-\end{figure}
 
@@
 
+\section{Old / Superceded}
+
+\subsection{Data Structures and Functions}
+
+Our program is used for image processing; thus, many computations are made on
+images, which are represented as arrays in \matlab\@. Additionally, we deal with
+matrices of images, so we need to model multi-dimensional matrices. Since
+multi-dimensional matrices can have arbitrarily many dimensions, we decided the
+best abstraction for this in Coq would be built on top of the \textsf{Vector}
+type which was originally implemented in the \colorlib\
+library~\cite{BLANQUI_2011} and eventually ported to the Coq standard library.
+In Coq, \textsf{Vector} is a dependent type~\cite{Bove2009,Thorsten_2010}; the
+number of elements in a \textsf{Vector} is part of its type, written
+\texttt{Vector.t A n} for a vector containing \(n\) elements of type \(A\).
+Since \textsf{Vector}s are polymorphic, we noted that a \textsf{Vector} could
+contain \textsf{Vector}s provided they are all the same type (which includes
+length, as noted). That is, we can stack \textsf{Vector}s to form, \emph{e.g.},
+\texttt{Vector.t (Vector.t A m) n}. From there, we defined a type-constructor
+\textsf{Matrix} taking a list of dimensions, which represents a \textsf{Vector}
+of \textsf{Vector}s. It can be used as the type of an arbitrary
+\(n\)-dimensional matrix containing elements of type \(A\).
+
+Due to the implementation of \textsf{Vector} in the Coq standard library, we
+needed to greatly extend the functionality of our \textsf{Matrix} type. To that
+end, we have begun implementing \textsf{Matrix} functions such as element
+access, which is non-trivial for matrices with an arbitrary number of
+dimensions\footnote{In particular because of the dependent types; see also
+\url{https://stackoverflow.com/q/66863226/4400820}}. Additionally, the existing
+implementation of element access in Coq's \textsf{Vector} type uses the
+unfamiliar \textsf{Fin} type, an abstraction for carrying proofs that a
+particular number belongs to finite range of numbers. In order to access an
+element of a \textsf{Vector}, one must supply such a proof; for our purposes, we
+use \texttt{Fin.of\_nat}, which gives either such a proof or a witness that the
+number is outside the range. This behavior is captured by the \textsf{sumbool}
+built-in, which required further research on our behalf. Given the number of
+ways that matrices can be indexed in \matlab\ and the complexity introduced by
+Coq's implementation of \textsf{Vector}s, we see this as the main bottleneck for
+our progress. We still need to implement many \textsf{Matrix} operations, such
+as accesses over ranges and over an entire dimension.
+
+\subsection{Binding Operations and Semantics}
+
+While we have not yet bound any of our functions or data types to the statements
+or expressions in \S~\ref{S:mmatlab}, we have naturally been thinking about them
+throughout the whole process. The functions we write are intended to implement
+the evaluation of the corresponding expressions. In the style of
+\texttt{Imp}~\cite{Pierce:SF1}, we intend to write a deterministic expression
+evaluation function; we will then provide semantics for program execution.
+Execution cannot be defined as a function in Coq because it may be
+non-terminating. We have had to consider what operations are be allowed and
+whether checks for undesired behavior (\emph{e.g.}, dynamic type mismatches)
+should be implemented in the underlying functions or in a semantic check.
+
+\begin{figure}[!t]
+    \textbf{Input}: image \(key\_art\), images \(thumbnails\), integer \(size\) \\
+    \textbf{Output}: image \(tiles\)
+    \begin{algorithmic}
+        \STATE{\(tiles \gets \textrm{MAT2TILES}(key\_art, size, size)\)}
+        \FOR{\(tile \in tiles\)}
+            \FOR{\(thumbnail, i \in thumbnails, [1..|thumbnails|]\)}
+                \STATE{\(mses_i \gets \textrm{immse}(tile, thumbnail)\)}
+            \ENDFOR
+            \STATE{\(best\_indices \gets \textrm{find} (mses = \min mses)\)}
+            \STATE{\(best\_index \gets best\_indices_0\)}
+            \STATE{\(tile \gets thumbnails_{best\_index}\)}
+        \ENDFOR
+    \RETURN{\(tiles\)}
+    \end{algorithmic}
+    \caption{Algorithm: Version 1}\label{alg:v1}
+\end{figure}
+
+\begin{figure}[!t]
+    \textbf{Input}: image \(key\_art\), images \(thumbnails\), integer \(size\) \\
+    \textbf{Output}: image \(tiles\)
+    \begin{algorithmic}
+        \STATE{\(tiles \gets \textrm{MAT2TILES}(key\_art, size, size)\)}
+        \STATE{\(used \gets \emptyset\)}
+        \STATE{\(threshold \gets 1\)}
+        \IF{\(|thumbnails| < |tiles|\)}
+            \STATE{\(threshold \gets \lceil{|tiles| / |thumbnails|}\rceil\)}
+        \ENDIF
+        \FOR{\(tile \in tiles\)}
+            \FOR{\(thumbnail, i \in thumbnails, [1..|thumbnails|]\)}
+                \STATE{\(mses_i \gets \textrm{immse}(tile, thumbnail)\)}
+            \ENDFOR
+            \STATE{\(best\_indices \gets \textrm{find} (mses = \min mses)\)}
+            \STATE{\(best\_index \gets best\_indices_0\)}
+            \WHILE{\(used_{best\_index} \ge threshold\)}
+                \STATE{\(mses_{best\_index} \gets \infty\)}
+                \STATE{\(best\_indices \gets \textrm{find} (mses = \min mses)\)}
+                \STATE{\(best\_index \gets best\_indices_0\)}
+            \ENDWHILE
+            \STATE{increment \(used_{best\_index}\)}
+            \STATE{\(tile \gets thumbnails_{best\_index}\)}
+        \ENDFOR
+        \RETURN{\(tiles\)}
+    \end{algorithmic}
+    \caption{Algorithm: Version 2}\label{alg:v2}
+\end{figure}
+
+\begin{figure}[!t]
+    \textbf{Input}: image \(key\_art\), images \(thumbnails\), integer \(size\) \\
+    \textbf{Output}: image \(tiles\)
+    \begin{algorithmic}
+        \STATE{}\COMMENT{Repeat Version 2}
+        \STATE{\(tiles \gets \textrm{Version2}(key\_art, thumbnails, size)\)}
+        \STATE{\(tiles \gets \textrm{imgaussfilt}(tiles)\)}
+        \STATE{\(strip\_size \gets \lfloor{size / 8}\rfloor\)}
+        \STATE{}\COMMENT{Blur by averaging}
+        \FOR{horizontal \(strip \in tiles\)}
+            \STATE{\(strip \gets \textrm{mean}({strip})\)}
+        \ENDFOR
+        \FOR{vertical \(strip \in tiles\)}
+            \STATE{\(strip \gets \textrm{mean}({strip})\)}
+        \ENDFOR
+        \RETURN{\(tiles\)}
+    \end{algorithmic}
+    \caption{Algorithm: Version 3}\label{alg:v3}
+\end{figure}
+
+\section{Goals and Timeline}
+
+As we neglected to include them in our original proposal, we now propose the
+following goals, taking into consideration the progress that we have made so far
+with the project:
+
+\begin{enumerate}
+
+    \item \textbf{Minimum Viable Product}: Successful modelling of a majority of
+        the desired features corresponding to matrix operations in \mmatlab\@.
+        Given the lack of multi-dimensional matrix libraries in Coq and the
+        necessary additions of several features in the Coq standard library,
+        this alone would be a notable achievement for us.
+
+    \item \textbf{Desired Goal}: Successful binding of our matrix operations to
+        \mmatlab\ statements and expressions, including the ability to evaluate
+        programs written in \mmatlab\@. A proof of termination, or sketch of
+        one, for our V1 algorithm is provided.
+
+    \item \textbf{Stretch Goal}: All of the above, and the addition of all
+        operations used in V2 and V3 algorithms (most notably, for loops that
+        increment across a range at varying intervals). Proofs that use
+        these features.
+
+\end{enumerate}
+
+Given what we have encountered so far, we propose the following modified
+timeline for project completion.
+
+\begin{itemize}
+    \item \textbf{03/31/21}: Progress Report Due
+    \item \textbf{04/09/21}: Significant progress made on most matrix operations
+        and data structures.
+    \item \textbf{04/16/21}: Binding \mmatlab\ syntax to operations written
+        in Coq; definition of semantics and program evaluator; formal proof
+        specification written.
+    \item \textbf{04/24/21}: Proof(s) should be written or scope adjusted
+        accordingly. All logic should be implemented; edits for readability
+        should be the main changes henceforth.
+    \item \textbf{04/26/21}: Final Proof Due
+    \item \textbf{05/03/21}: Final Paper Due
+    \item \textbf{05/05/21}: Final Presentation Delivered
+\end{itemize}
+
+We provide the following ``diff'' between the scope of our original project
+proposal and our new proposal:
+\begin{itemize}
+    \item Heightened emphasis on implementation of matrix operations, since they
+        are proving to be non-trivial.
+    \item Addition of goal tiers.
+    \item Discussion on Hoare logic.
+\end{itemize}
+
 {\printbibliography}

### 2021-05-02 paper/paper.tex
paper: proof-reading pass

@@
 
+\begin{figure}[htp]
+    \textbf{Input}: image \(key\_art\), images \(thumbnails\), integer \(size\) \\
+    \textbf{Output}: image \(tiles\)
+    \begin{algorithmic}
+        \STATE{\(tiles \gets \textrm{MAT2TILES}(key\_art, size, size)\)}
+        \FOR{\(tile \in tiles\)}
+            \FOR{\(thumbnail, i \in thumbnails, [1..|thumbnails| ]\)}
+                \STATE{\(mses_i \gets \textrm{immse}(tile, thumbnail)\)}
+            \ENDFOR
+            \STATE{\(best\_indices \gets \textrm{find} (mses = \min mses)\)}
+            \STATE{\(best\_index \gets best\_indices_0\)}
+            \STATE{\(tile \gets thumbnails_{best\_index}\)}
+        \ENDFOR
+    \RETURN{\(tiles\)}
+    \end{algorithmic}
+    \caption{Algorithm: Version 1}\label{alg:v1}
+\end{figure}
+
+\begin{figure}[htp]
+    \textbf{Input}: image \(key\_art\), images \(thumbnails\), integer \(size\) \\
+    \textbf{Output}: image \(tiles\)
+    \begin{algorithmic}
+        \STATE{\(tiles \gets \textrm{MAT2TILES}(key\_art, size, size)\)}
+        \STATE{\(used \gets \emptyset\)}
+        \STATE{\(threshold \gets 1\)}
+        \IF{\(|thumbnails| < |tiles|\)}
+            \STATE{\(threshold \gets \lceil{|tiles| / |thumbnails|}\rceil\)}
+        \ENDIF
+        \FOR{\(tile \in tiles\)}
+            \FOR{\(thumbnail, i \in thumbnails, [1..|thumbnails| ]\)}
+                \STATE{\(mses_i \gets \textrm{immse}(tile, thumbnail)\)}
+            \ENDFOR
+            \STATE{\(best\_indices \gets \textrm{find} (mses = \min mses)\)}
+            \STATE{\(best\_index \gets best\_indices_0\)}
+            \WHILE{\(used_{best\_index} \ge threshold\)}
+                \STATE{\(mses_{best\_index} \gets \infty\)}
+                \STATE{\(best\_indices \gets \textrm{find} (mses = \min mses)\)}
+                \STATE{\(best\_index \gets best\_indices_0\)}
+            \ENDWHILE
+            \STATE{increment \(used_{best\_index}\)}
+            \STATE{\(tile \gets thumbnails_{best\_index}\)}
+        \ENDFOR
+        \RETURN{\(tiles\)}
+    \end{algorithmic}
+    \caption{Algorithm: Version 2}\label{alg:v2}
+\end{figure}
+
+\begin{figure}[htp]
+    \textbf{Input}: image \(key\_art\), images \(thumbnails\), integer \(size\) \\
+    \textbf{Output}: image \(tiles\)
+    \begin{algorithmic}
+        \STATE{}\COMMENT{Repeat Version 2}
+        \STATE{\(tiles \gets \textrm{Version2}(key\_art, thumbnails, size)\)}
+        \STATE{\(tiles \gets \textrm{imgaussfilt}(tiles)\)}
+        \STATE{\(strip\_size \gets \lfloor{size / 8}\rfloor\)}
+        \STATE{}\COMMENT{Blur by averaging}
+        \FOR{horizontal \(strip \in tiles\)}
+            \STATE{\(strip \gets \textrm{mean}({strip})\)}
+        \ENDFOR
+        \FOR{vertical \(strip \in tiles\)}
+            \STATE{\(strip \gets \textrm{mean}({strip})\)}
+        \ENDFOR
+        \RETURN{\(tiles\)}
+    \end{algorithmic}
+    \caption{Algorithm: Version 3}\label{alg:v3}
+\end{figure}
+
 Our original goal was to examine these algorithms as implemented in the \matlab\
@@
 \section{Problem Description}
-Many non-trivial programs make use of arrays, and our \texttt{zelda-mosaic} program is no different.
-Because our goal was to prove the termination of the program when given a well-formed input and the core data structure modified throughout the program was an image (represented as an array), it is helpful to first describe the array operations used and the invariants of an array that contribute to ``well-formed-ness".
+
+Many non-trivial programs make use of arrays, and our \texttt{zelda-mosaic}
+program is no different. Because our goal was to prove the termination of the
+program when given a well-formed input and the core data structure modified
+throughout the program was an image (represented as an array), it is helpful to
+first describe the array operations used and the invariants of an array that
+contribute to the \emph{well-formed} property.
 
 \subsection{Array Usage}
-\matlab's support for multi-dimensional arrays makes it amenable to image-processing tasks such as ours.
-The algorithms we wrote are described at high-level in pseudocode in Figures \figurename~\ref{alg:v1}, \ref{alg:v2}, and \ref{alg:v3}, but here we explain them in layman's terms to help motivate some features of our proof.\\
-
-As described in our pseudocode, the inputs to our program were a piece of \textit{key art} to be recomposed with \textit{thumbnails} of a certain \textit{size}, defined as the width and height of the thumbnails (all have been coerced to be square-shaped).
-The original representation of \textit{key art} in \matlab is a two-dimensional array (with indices corresponding to a particular pixel's row and column) containing an additional array of three values - one each for the red, green, and blue color channels.
-So far, this is an array of three dimensions.
-We then divide this image into \textit{chunks}, where each \textit{chunk} will be replaced by a \textit{thumbnail} in the final product.
-For \textit{key art} of dimensions $m \times n$ (with $m$ corresponding to and $n$ corresponding to columns), each image will be divided into $(m / size) \times (n / size)$ \textit{chunks}, which will each be replaced by a \textit{thumbnail} as appropriate.
-These \textit{chunks}, each of size $size \times size$, are then stored in a two-dimensional array, with each \textit{chunk} retaining its original placement in the \textit{key art}.
-As a result we now have a five-dimensional array, and in order to access a particular value from a color channel, we need the following values:
+
+\matlab's support for multi-dimensional arrays makes it amenable to
+image-processing tasks such as ours. The algorithms we wrote are described at
+high-level in pseudocode in \figurename{s}~\ref{alg:v1},~\ref{alg:v2},
+and~\ref{alg:v3}, but here we explain them in layman's terms to help motivate
+some features of our proof.
+
+As described in our pseudocode, the inputs to our program were a piece of
+\emph{key art} to be recomposed with \emph{thumbnails} of a certain \emph{size},
+which is the width and height of the square thumbnails. The original
+representation of the key art in \matlab\ is a two-dimensional array (with
+indices corresponding to a particular pixel's row and column) containing an
+additional array of three values---one each for the red, green, and blue color
+channels. In total, the array has three dimensions. We then divide this image
+into \emph{chunks}, where each chunk will be replaced by a
+thumbnail in the final product. For key art of dimensions $m
+\times n$ (with $m$ corresponding to rows and $n$ corresponding to columns), each
+image will be divided into $(m / \var{size}) \times (n / \var{size})$ chunks, which
+will each be replaced by a thumbnail as appropriate. These chunks,
+each of size $\var{size} \times \var{size}$, are then stored in a two-dimensional array,
+with each chunk retaining its original placement in the key art.
+As a result we now have a five-dimensional array, and in order to access a
+particular value from a color channel, we need the following values:
 
 \begin{itemize}
-\item the \textit{chunk row} it is in
-\item the \textit{chunk column} it is in
-\item the \textit{x-offset} of the pixel into its chunk
-\item the \textit{y-offset} of the pixel into its chunk
-\item the \textit{color channel} desired out of red, green, and blue
+\item the \emph{chunk row} it is in
+\item the \emph{chunk column} it is in
+\item the \emph{x offset} of the pixel into its chunk
+\item the \emph{y offset} of the pixel into its chunk
+\item the \emph{color channel} desired out of red, green, and blue
 \end{itemize}
 
-Thus, once a piece of \textit{key art} is successfully divided into \textit{chunks}, the image is effectively represented as an array whose dimensions correspond to the above values.
-The \textit{chunks} are iterated through in row-major order, gradually being replaced in-place by \textit{thumbnails}.
-Consequently, the state of our core array maintains some matrix-like properties that are supported by \matlab\@ arrays, but not particular to \matlab\@ in general:
+Thus, once a piece of key art is successfully divided into chunks, the image is
+effectively represented as an array whose dimensions correspond to the above
+values. The chunks are iterated through in row-major order, gradually being
+replaced in-place by thumbnails. Consequently, the state of our core array
+maintains some matrix-like properties that are supported by \matlab\ arrays,
+but not particular to \matlab\ in general:
 
 \begin{itemize}
-    \item The shape of a matrix \textbf{$M$} can be described by a list of \textit{d} dimensions, where each dimension's value $v_i$ refers to the number of elements held along that dimension.
-For example, an array with dimensions $3 \times 4 \times 2$ contains 3 arrays, each of which contains 4 arrays of 2 elements.
-$v_1 = 3$, $v_2 = 4$, and $v_3 = 2$.
-    \item Once the shape of \textbf{$M$} is defined and all of the elements have been populated, the shape will not change and all of the elements will remain populated.
-For example, the aforementioned $3 \times 4 \times 2$ array must always contain 3 arrays, each of which contains 4 arrays of 2 elements.
-    \item Each $i$-th dimension from $1$ to $d-1$ must contain $v_i$ elements, each of which is an array of length $v_{i+1}$; in other words, the length of each element in a particular dimension must be identical to one another.
+    \item The shape of a matrix \(M\) can be described by a list of \(d\)
+        dimensions, where each dimension's value \(v_i\) refers to the number of
+        elements held along that dimension. For example, an array with
+        dimensions \(3 \times 4 \times 2\) contains 3 arrays, each of which
+        contains 4 arrays of 2 elements. The dimensions list contains \(v_1 =
+        3\), \(v_2 = 4\), and \(v_3 = 2\).
+    \item Once the shape of \(M\) is defined and all of the elements have been
+        populated, the shape will not change and all of the elements will remain
+        populated. For example, the aforementioned \(3 \times 4 \times 2\) array
+        must always contain 3 arrays, each of which contains 4 arrays of 2
+        elements.
+    \item Each \(i\)-the dimension from \(1\) to \(d-1\) must contain \(v_i\)
+        elements, each of which is an array of length \(v_{i+1}\); in other
+        words, the length of each element in a particular dimension must be
+        identical to one another.
 \end{itemize}
 
-In summary, these properties enforce a notion of ``well-formed-ness" that ensures that the size $v_i$ of each dimension remains fixed and that all element in a particular dimension $i$ have the same length as one another.
-This property is generally true of matrices, hence the references to both an ``arrays" and ``matrices" in this section.
-Unless otherwise noted, ``array" refers to the generic \matlab\@ object and ``matrix" refers to an ``array" with these properties when discussed in the scope of this paper.
-
-\subsection{Operations} \matlab\@ supports a number of operations that can be used to manipulate arrays, but only a subset are used in our program.
-To illustrate the operations used, we begin with the following visualization of a three-dimensional array \textbf{$E$} with dimensions $3 \times 4 \times 2$, as it is used in \matlab:\\
+In summary, these properties describe the notion of a ``well-formed'' matrix
+that ensures that the size \(v_i\) of each dimension remains fixed and that all
+elements in a particular dimension \(i\) have the same length as one another.
+This property is generally true of matrices, hence the references to both an
+``arrays'' and ``matrices'' in this section. Unless otherwise noted, ``array''
+refers to the generic \matlab\ object and ``matrix'' refers to an ``array'' with
+these properties when discussed in the scope of this paper.
+
+\subsection{Operations}
+
+\matlab\ supports a number of operations that can be used to manipulate arrays,
+but only a subset are used in our program. To illustrate the operations used, we
+begin with a visualization of a three-dimensional array \(E\) with dimensions
+\(3 \times 4 \times 2\) below and discuss how it is used in \matlab.
 
@@
 
-\begin{figure}[h!]
-    \centering
+\begin{center}
     \begin{tikzpicture}[every node/.style={anchor=north east,fill=white, minimum width=.7cm}]
@@
     \end{tikzpicture}
-
-\end{figure}
+\end{center}
 
 \subsubsection{Element Access}
-For the matrix \textbf{$E$}, each element would be indexed according to the following mapping.
-Under this scheme, for example, the element $8$ in \textbf{$E$} would have the index $(1, 4, 2)$.
-This indexing scheme is the general pattern used to access one particular element out of a multi-dimensional array.
-Note that arrays in \matlab\@ have indices starting at 1.
-
-\begin{figure}[h!]
-    \centering
+
+For the matrix \(E\), each element would be indexed according to the following
+mapping. Under this scheme, for example, the element \(8\) in \(E\) would
+have the index \((1, 4, 2)\). This indexing scheme is the general pattern used to
+access one particular element out of a multi-dimensional array. Note that arrays
+in \matlab\ have indices starting at 1. The indexing scheme is demonstrated
+below.
+
+\begin{center}
     \begin{tikzpicture}
@@
     \end{tikzpicture}
-\end{figure}
-
-\subsubsection{Linearized Indexing} \label{linearizedindexing}
-Note that in $E$, elements are ordered in ascending order with respect to their indexing.
-First, sub-array $(1)$ is populated, and $(1, 1)$ is the first array within it, so it is filled with elements $1$ and $2$.
-Array $(1, 2)$ is then populated by elements $3$ and $4$.
-After all four of \textbf{$E$}'s sub-arrays have been populated, sub-array $(2)$ is populated, and so on until the array is full.
-
-Given the numbering scheme in $E$, one can imagine a ``flattened" version of $E$ that is simply an ordered one-dimensional array containing all of $E$'s elements; simply put, an array containing the elements $1$ to $24$, in numerical order.
-\matlab\@ actually allows arrays of arbitrarily many dimensions to be logically ``flattened" during accesses so that elements can be accessed with one index, like in a one-dimensional array.
-We found this property to be somewhat interesting and challenging to model, and we refer to the ``flattening" of a matrix as ``linearization".
-
-The process of linearizing a matrix is perhaps
-better explained with a two-dimensional example. Imagine a general two-dimensional $3 \times 4$ matrix, such as the one below, where:
+\end{center}
+
+\subsubsection{Linearized Indexing}\label{S:linearized_indexing}
+
+Note that in \(E\), elements are ordered in ascending order with respect to
+their indexing. First, sub-array \((1)\) is populated, and \((1, 1)\) is the
+first array within it, so it is filled with elements \(1\) and \(2\). Array
+\((1, 2)\) is then populated by elements \(3\) and \(4\). After all four of
+\(E\)'s sub-arrays have been populated, sub-array \((2)\) is populated, and so
+on until the array is full.
+
+Given the numbering scheme in \(E\), one can imagine a ``flattened'' version of
+\(E\) that is an one-dimensional array containing all of \(E\)'s elements;
+simply put, an array containing the elements \(1\) to \(24\) in numerical
+order. \matlab\ actually allows arrays of arbitrarily many dimensions to be
+logically flattened during accesses so that elements can be accessed with one
+index, like in a one-dimensional array. We found this property to be somewhat
+interesting and challenging to model, and we refer to the flattening of a matrix
+as ``linearization.''
+
+The process of linearizing a matrix is perhaps better explained with a
+two-dimensional example. Imagine a general two-dimensional \(3 \times 4\) matrix,
+such as the one below, where:
 \begin{equation*}
@@
 
-Under this scheme, any element in a matrix can be accessed via its position in the linearized version of that matrix.
-For example, the element $12$ in \textbf{$E$} can be accessed by the index $(12)$ (note that 12 refers to the index, not the elements) and in the previous example, the element $f$ can be accessed by the index $(6)$.
-
-\subsubsection{Ranged Indexing} \label{rangedindexing}
-Broadly speaking, \matlab\@ supports the notion of a range of numbers specified by a starting number and a stopping number, which includes every number between the starting number and the stopping number (inclusive).
-This abstraction can be used to access or assign elements within a sub-matrix of a particular matrix.
-For example, suppose in matrix $E$ you only cared about the first matrix in the first dimension, the second/third/fourth columns within the second dimension, and the second row within the third dimension.
-This sub-matrix can be specified by the index $(1, 2:4, 2)$ and would return the matrix:
+Under this scheme, any element in a matrix can be accessed via its position in
+the linearized version of that matrix. For example, the element \(12\) in \(E\)
+can be accessed by the index \((12)\) and in the previous example, the element
+\(f\) can be accessed by the index \((6)\).
+
+\subsubsection{Ranged Indexing}\label{S:ranged_indexing}
+
+Broadly speaking, \matlab\ supports the notion of a range of numbers specified
+by a starting index and a stopping index, which includes every index between the
+starting index and the stopping index (inclusive). This abstraction can be used
+to access or assign elements within a sub-matrix of a particular matrix. For
+example, suppose in matrix \(E\) you only cared about the first matrix in the
+first dimension; the second, third, and fourth columns within the second
+dimension; and the second row within the third dimension. This sub-matrix can be
+specified by the index \((1, 2:4, 2)\) and would be the matrix
+
 \begin{equation*}
@@
         4 & 6 & 8
-    \end{bmatrix}
+    \end{bmatrix}.
 \end{equation*}
 
-Suppose that you wanted to get the second/third/fourth columns within the second dimension and second row within the third dimension for \textit{all three} of the outer-most dimension of \textbf{$E$} but did not know a priori how many elements it had.
-\matlab\@ supports an ``inclusive range" operator, represented by a colon \textbf{:}, which is taken to mean the range encompassing all values in a particular dimension.
-The sub-matrix described here would be specified by the index $(:, 2:4, 2)$ and represents the following matrix:
-
-\begin{figure}[h!]
-    \centering
+Suppose that you wanted to get the second, third, and fourth columns within the
+second dimension and the second row within the third dimension for \emph{all
+three} of the outer-most dimension of \(E\) but did not know \textit{a priori}
+how many elements it had. \matlab\ supports an ``inclusive range'' operator,
+represented by a colon \texttt{:}. This operator represents the range
+encompassing all values in a particular dimension. The sub-matrix described here
+would be specified by the index $(:, 2:4, 2)$ and represents the following
+matrix:
+
+\begin{center}
     \begin{tikzpicture}[every node/.style={anchor=north east,fill=white, minimum width=.7cm}]
@@
     \end{tikzpicture}
-
-\end{figure}
-
-\subsubsection{Matrix Operations} \label{matrixoperations}
-Details about \matlab\@ arrays pertaining to indexing are perhaps some of their most interesting features, but we also use some typical matrix operators in our program.
-Generally speaking, matrix operators can be generalized to any number of dimensions, as long as the dimensions of the arrays agree for the operation in question.
-There are two broad classes of matrix operators that we make use of:
-
-\begin{itemize}
-    \item \textbf{Element-wise}: For two matrices of the same dimensionality, the operator is applied on each of the two elements containing the same index in the two matrices.
-    \item \textbf{All Elements}: For an array and a scalar, the operation and scalar (e.g. add 1) is applied to every element in the matrix.
-\end{itemize}
+\end{center}
+
+\subsubsection{Matrix Operations}\label{S:matrix_operations}
+
+Details about \matlab\ arrays pertaining to indexing are perhaps some of their
+most interesting features, but we also use some typical matrix operators in our
+program. Generally speaking, matrix operators can be generalized to any number
+of dimensions, as long as the dimensions of the arrays agree for the operation
+in question. There are two broad classes of matrix operators that we make use
+of:
+
+\begin{description}
+    \item[Element-wise] For two matrices of the same dimensionality, the
+        operator is applied on each of the two elements containing the same
+        index in the two matrices.
+    \item[All Elements] For an array and a scalar, the operation and scalar
+        (\textit{e.g.}, add 1) is applied to every element in the matrix.
+\end{description}
 
 \subsection{Modeling the Problem}
-Aside from the core matrix operations, we anticipated that we would need to model the state of our program at all points of execution since our goal was to prove termination.
-To this end, we would need to identify all of the syntax that we use in our program and specify their associated semantics.
-The upshot of modeling our program in this way would be the ability to remains to prove its termination of our given that certain preconditions hold.
-
-Hoare logic~\cite{Hoare_1969} is a type of logic that is often used to solve problems of this class.
-Partial Hoare logic is a weak statement about programs; namely, given preconditions \(P\) and postconditions \(Q\) and a program \(C\), the Hoare triple \(\{P\}C\{Q\}\) asserts that \emph{if} \(C\) terminates, then the pre-conditions imply the post-conditions.
-Total Hoare logic, by contrast, is a stronger version of this claim: \(C\) is asserted to terminate.
-Thus a proof in total Hoare logic necessitates a proof that \(C\) terminates; the inference rules are accordingly adjusted.
-The primary adjustment is to the rule for while
+
+Aside from the core matrix operations, we anticipated that we would need to
+model the state of our program at all points of execution since our goal was to
+prove termination. To this end, we would need to identify all of the syntax that
+we use in our program and specify their associated semantics. The upshot of
+modeling our program in this way would be the ability to remains to prove its
+termination of our given that certain preconditions hold.
+
+Hoare logic~\cite{Hoare_1969} is a type of logic that is often used to solve
+problems of this class. Partial Hoare logic is a weak statement about programs;
+namely, given preconditions \(P\) and postconditions \(Q\) and a program \(C\),
+the Hoare triple \(\{P\}C\{Q\}\) asserts that \emph{if} \(C\) terminates, then
+the pre-conditions imply the post-conditions. Total Hoare logic, by contrast, is
+a stronger version of this claim: \(C\) is asserted to terminate. Thus a proof
+in total Hoare logic necessitates a proof that \(C\) terminates; the inference
+rules are accordingly adjusted. The primary adjustment is to the rule for while
 loops, which are guaranteed to terminate by relation to a decreasing sequence
@@
 
-When we started this project, we had in mind that the preconditions would be statements about the program's inputs; for example, ``the length and width of \textit{key art} are always multiples of the \textit{thumbnails'} size".
-Ultimately, we did not get far enough in modeling our program to devise a set of pre-conditions, but the point remains that we needed to be able to model the control flow of our entire program in order to do this.
-
-\section{Design / Approach}
-
-We first set out to define the syntax and semantics of the subset of \matlab\ used in our program, making the assumption that once we finished this it would be somewhat trivial to bind the defined syntax to behavior in Coq.
-Ultimately, the majority of our effort in this project ended going towards the implementation of our multi-dimensional matrix operations, which was made difficult by the number of properties that our matrices uphold as well as our desire to write generic Coq functions that would work on matrices of arbitrarily many dimensions.
-We spent a while working with Coq's built-in \textsf{Vector} type before coming to the conclusion that its dependent typing caused difficulties in behavior specification.
-To this end, we ended up implementing our own data types to describe the state of matrices in our program; this task was unexpectedly challenging and once we began working on it, it became the only thing we worked on for the remainder of the project.
+When we started this project, we had in mind that the preconditions would be
+statements about the program's inputs; for example, ``the length and width of
+key art are always multiples of the thumbnails' size.'' Ultimately, we did not
+get far enough in modeling our program to devise a set of pre-conditions, but
+the point remains that we needed to be able to model the control flow of our
+entire program in order to do this.
+
+\section{Design and Approach}\label{S:design}
+
+We first set out to define the syntax and semantics of the subset of \matlab\
+used in our program, making the assumption that once we finished this it would
+be somewhat trivial to bind the defined syntax to behavior in Coq. Ultimately,
+the majority of our effort in this project ended going towards the
+implementation of our multi-dimensional matrix operations, which was made
+difficult by the number of properties that our matrices uphold as well as our
+desire to write generic Coq functions that would work on matrices of arbitrarily
+many dimensions. We spent a while working with Coq's built-in \textsf{Vector}
+type before coming to the conclusion that its dependent typing caused
+difficulties in behavior specification. To this end, we ended up implementing
+our own data types to describe the state of matrices in our program; this task
+was unexpectedly challenging and once we began working on it, it became the only
+thing we worked on for the remainder of the project.
 
@@
 definitions are elided here. However, we wish to draw attention to some of the
-less apparent and more interesting syntax present in \mmatlab\@:
+less apparent and more interesting syntax present in \mmatlab:
 
@@
         matrix, is a matrix with all elements incremented by the number. Similar
-        behaviors hold for \textsf{MultExp} and \textsf{DivExp}. See Section \ref{matrixoperations} for the distinction between these operators.
+        behaviors hold for \textsf{MultExp} and \textsf{DivExp}.
 
@@
         starting number and stopping number may be provided, generating an
-        inclusive list of all numbers between them. This is discussed in Section \ref{rangedindexing}
-
-    \item \textsf{IndexExp}: In support of the RangeExp expression, the syntax to access elements from an array is overloaded in \matlab\@. They
+        inclusive list of all numbers between them. This is discussed in
+        \S~\ref{S:ranged_indexing}.
+
+    \item \textsf{IndexExp}: This expression pairs a variable with an indexer,
+        which can be an almost arbitrary expression. \matlab\ does restrict the
+        syntactic placement of certain indexing operations.
 
@@
 
-As mentioned in \S~\ref{S:design_dependent}, our original formulation of
-matrices as nested dependent \texttt{Vector.t}s quickly grew in complexity.
-We abandoned that approach in favor of a type that is not guaranteed to behave
-well, but whose definition makes defining operations more
-natural~\cite{SO_2021_2}. This approach removes the proofs that dependent types
-provide automatically, so we have to give those ourselves.
+As mentioned in \S~\ref{S:design}, our original formulation of matrices as
+nested dependent \texttt{Vector.t}s quickly grew in complexity. We abandoned
+that approach in favor of a type that is not guaranteed to behave well, but
+whose definition makes defining operations more natural~\cite{SO_2021_2}. This
+approach removes the proofs that dependent types provide automatically, so we
+have to give those ourselves.
 
@@
 
-\paragraph{Linearizing a matrix} See \S~\ref{S:design_linearize} for more
+\paragraph{Linearizing a matrix} See \S~\ref{S:linearized_indexing} for more
 details on matrix linearization. To define linearization in Coq, we use a
@@
 \paragraph{Linearly indexing a matrix} Given the operation to linearise a
-matrix, one-based linearly indexing is easy. See \S~\ref{S:design_linear_index}
+matrix, one-based linearly indexing is easy. See \S~\ref{S:linearized_indexing}
 for details on linear indexing. The function \(\func{nth}\) takes a matrix \(m\)
@@
 \(\func{None}\)---a simple \texttt{if} or \texttt{match} guard would cover this
-case.
+case. We note that our linear indexing supports only a single index, while
+\matlab\ allows any index prior to the last to start linearization.
 
@@
 
-\begin{figure}[!t]
-    \textbf{Input}: image \(key\_art\), images \(thumbnails\), integer \(size\) \\
-    \textbf{Output}: image \(tiles\)
-    \begin{algorithmic}
-        \STATE{\(tiles \gets \textrm{MAT2TILES}(key\_art, size, size)\)}
-        \FOR{\(tile \in tiles\)}
-            \FOR{\(thumbnail, i \in thumbnails, [1..|thumbnails|]\)}
-                \STATE{\(mses_i \gets \textrm{immse}(tile, thumbnail)\)}
-            \ENDFOR
-            \STATE{\(best\_indices \gets \textrm{find} (mses = \min mses)\)}
-            \STATE{\(best\_index \gets best\_indices_0\)}
-            \STATE{\(tile \gets thumbnails_{best\_index}\)}
-        \ENDFOR
-    \RETURN{\(tiles\)}
-    \end{algorithmic}
-    \caption{Algorithm: Version 1}\label{alg:v1}
-\end{figure}
-
-\begin{figure}[!t]
-    \textbf{Input}: image \(key\_art\), images \(thumbnails\), integer \(size\) \\
-    \textbf{Output}: image \(tiles\)
-    \begin{algorithmic}
-        \STATE{\(tiles \gets \textrm{MAT2TILES}(key\_art, size, size)\)}
-        \STATE{\(used \gets \emptyset\)}
-        \STATE{\(threshold \gets 1\)}
-        \IF{\(|thumbnails| < |tiles|\)}
-            \STATE{\(threshold \gets \lceil{|tiles| / |thumbnails|}\rceil\)}
-        \ENDIF
-        \FOR{\(tile \in tiles\)}
-            \FOR{\(thumbnail, i \in thumbnails, [1..|thumbnails|]\)}
-                \STATE{\(mses_i \gets \textrm{immse}(tile, thumbnail)\)}
-            \ENDFOR
-            \STATE{\(best\_indices \gets \textrm{find} (mses = \min mses)\)}
-            \STATE{\(best\_index \gets best\_indices_0\)}
-            \WHILE{\(used_{best\_index} \ge threshold\)}
-                \STATE{\(mses_{best\_index} \gets \infty\)}
-                \STATE{\(best\_indices \gets \textrm{find} (mses = \min mses)\)}
-                \STATE{\(best\_index \gets best\_indices_0\)}
-            \ENDWHILE
-            \STATE{increment \(used_{best\_index}\)}
-            \STATE{\(tile \gets thumbnails_{best\_index}\)}
-        \ENDFOR
-        \RETURN{\(tiles\)}
-    \end{algorithmic}
-    \caption{Algorithm: Version 2}\label{alg:v2}
-\end{figure}
-
-\begin{figure}[!t]
-    \textbf{Input}: image \(key\_art\), images \(thumbnails\), integer \(size\) \\
-    \textbf{Output}: image \(tiles\)
-    \begin{algorithmic}
-        \STATE{}\COMMENT{Repeat Version 2}
-        \STATE{\(tiles \gets \textrm{Version2}(key\_art, thumbnails, size)\)}
-        \STATE{\(tiles \gets \textrm{imgaussfilt}(tiles)\)}
-        \STATE{\(strip\_size \gets \lfloor{size / 8}\rfloor\)}
-        \STATE{}\COMMENT{Blur by averaging}
-        \FOR{horizontal \(strip \in tiles\)}
-            \STATE{\(strip \gets \textrm{mean}({strip})\)}
-        \ENDFOR
-        \FOR{vertical \(strip \in tiles\)}
-            \STATE{\(strip \gets \textrm{mean}({strip})\)}
-        \ENDFOR
-        \RETURN{\(tiles\)}
-    \end{algorithmic}
-    \caption{Algorithm: Version 3}\label{alg:v3}
-\end{figure}
-
 \section{Goals and Timeline}

### 2021-05-02 paper/paper.tex
paper: finish the theorems

@@
 
-\paragraph{Linearizing a matrix} See \S~\ref{S:linearized_indexing} for more
-details on matrix linearization. To define linearization in Coq, we use a
-two-stage unpacking as with other matrix operations. The function
-\(\func{linearize'}\) turns scalars into a singleton list and recursively
-flat-maps itself over matrices to produce the one-dimensional matrix as a list.
-The function \(\func{linearize}\) unpacks a matrix, then transforms the contents
-by wrapping the result of mapping \(\func{linearize'}\) into \(\func{Scalar}\)s
-in a \(\func{Matrix}\) and transforms the shape by taking the product of the
-(normalized) shape of the matrix. Recall that the length of the linearized
-vector is the product of the sizes of all the dimensions. The definitions are
-spelled out below:
+\paragraph{Linearizing a matrix}\label{S:linearize_coq} See
+\S~\ref{S:linearized_indexing} for more details on matrix linearization. To
+define linearization in Coq, we use a two-stage unpacking as with other matrix
+operations. The function \(\func{linearize'}\) turns scalars into a singleton
+list and recursively flat-maps itself over matrices to produce the
+one-dimensional matrix as a list. The function \(\func{linearize}\) unpacks a
+matrix, then transforms the contents by wrapping the result of mapping
+\(\func{linearize'}\) into \(\func{Scalar}\)s in a \(\func{Matrix}\) and
+transforms the shape by taking the product of the (normalized) shape of the
+matrix. Recall that the length of the linearized vector is the product of the
+sizes of all the dimensions. The definitions are spelled out below:
 
@@
     definition) with respect to some shape \(t\), then the following matrix is
-    also well-formed: \(\{ \var{shape} = |\var{ms}| :: t; \var{contents} =
-    \func{Matrix}\, \var{ms} \}\).
+    also well-formed in the inductive definition: \(\{ \var{shape} = |\var{ms}|
+    :: t; \var{contents} = \func{Matrix}\, \var{ms} \}\).
 \end{lem}
@@
 
-\begin{thm}[Linearize has resulting shape the product of the original]\label{Th:linearize_product}
+\begin{thm}[A linearized matrix has length the product of the shape]\label{Th:linearize_product}
+    For any well-formed matrix \(m\), the length of the linearized contents is
+    equal to the product of the computed shape of the matrix; that is,
+    \(|\func{linearize'}\, (\func{contents}\, m)| = \prod
+    \func{compute\_shape}\, (\func{contents}\, m)\).
 \end{thm}
 
+\begin{proof}
+    We proceed by induction on the evidence for the inductive variant of
+    the well-formed property (thanks to Theorem~\ref{Th:well_formed_agree}).
+
+    The cases for \(\iname{WF\_Empty}\) and \(\iname{WF\_Scalar}\) are
+    straightforward: computation via the definitions shows that both sides are
+    euqal.
+
+    In the inductive case for \(\iname{WF\_Cons}\), we combine a few facts about
+    flat-map, length, and list concatenation with the inductive hypothesis to
+    simplify the proof. The key piece of the proof is to show that the product
+    of the computed shape of some \(\func{Matrix}\, \var{ms}\) is equal to the
+    length of \(\var{ms}\) times the product of the computed shape of some \(m\)
+    (both pieces of the evidence from the inductive judgement, and both
+    with well-formed hypotheses). But this equality is straightforward by case
+    analysis on \(\var{ms}\). The empty case is trivial. In the non-empty case,
+    we know that the elements in the list and \(m\) have the same computed
+    shaped because they are well-formed with respect to the same shape (see
+    Lemma~\ref{Lem:wf_same_shape}). This is enough to finish.
+\end{proof}
+
 \begin{thm}[Linearize preserves the well-formed property]\label{Th:linearize_wf}
+    For any well-formed matrix \(m\), the matrix \(\func{linearize}\, m\) is
+    also well-formed.
 \end{thm}
+
+\begin{proof}
+    We proceed by induction on the contents of \(m\), with the usual case
+    analysis on the shape to eliminate contradictory cases.
+
+    The base case is trivial by the definition of the well-formed property and
+    the linearize function.
+
+    In the inductive case, there are two major statements to prove. Firstly, we
+    must prove that the length of the mapped \(\func{linearize'}\)d list is the
+    product of the computed shape. Secondly, we must prove that all elements in
+    the mapped \(\func{linearize'}\)d list are well-formed. See
+    \S~\ref{S:linearize_coq} for the definition of \(\func{linearize}\) that
+    leads to this mapped list.
+
+    For the first statement, we use the fact that map does not change the length
+    of the list together with Theorem~\ref{Th:linearize_product} to see that we
+    only need to prove that a particular matrix is well-formed. For that proof,
+    we combine Lemma~\ref{Lem:wfI_all_wf_t} with
+    Theorem~\ref{Th:well_formed_agree} and the well-formed hypothesis.
+
+    For the second statement, we only need to show that everything in the
+    mapped \(\func{linearize'}\)d list is a \(\func{Scalar}\)---but we are
+    mapping the \(\func{Scalar}\) constructor over the list, so this is of
+    course true.
+\end{proof}
 

### 2021-05-02 paper/paper.tex
paper: quick paragraph about range-based indexing

@@
 
+\paragraph{Range-based indexing} See \S~\ref{S:ranged_indexing} for details on
+range-based indexing. Our initial attempt, though incomplete, defines a type for
+the different types of range indexers and a function for indexing a sub-matrix
+via a list of range indexers. We also make use of an auxiliary data function to
+convert a list of options to an optional list, which we have proven
+correct\footnote{The result is \(\func{None}\) if there is a \(\func{None}\)
+anywhere in the list. The result is a list of values if-and-only-if the two
+lists have the same length and they agree on the values at all possible
+indices.}.
+
 \subsection{Key Theorems about Matrices}\label{S:matrix_thm}

### 2021-05-02 paper/paper.tex
paper: write about Coq

@@
 All code may be found at the project repository~\cite{zelda_mosaic_proof}. Most
-of the work in this section corresponds to the file \texttt{Lib Matrix.v}.
+of the work in this section corresponds to the file \texttt{LibMatrix.v}.
 
@@
 
+There are three major advanced Coq features that assisted our development of the
+matrix library. These are
+\begin{inlist}
+\item the use of Ltac to simplify and automate proofs;
+\item the use of hint databases to enhance the \texttt{auto} proof-search
+    tactic; and
+\item the use of specialized induction principles to enhance induction with
+    necessary information.
+\end{inlist}
+We cover each in turn.
+
+\subsubsection{Automating Coq via Ltac}\label{S:coq_ltac}
+
+Ltac is one of Coq's languages for defining custom tactics. It can be used for
+simple abbreviations (\textit{e.g.}, \texttt{gen} for \texttt{generalize dependent}) and
+complex proof machinery (\textit{e.g.}, find two equalities over the same
+variable that contradict each other and exploit the contradiction). We use Ltac
+to enhance several standard Coq details by use of the file \texttt{LibMatrix.v}
+from \textit{Programming Language Foundations}~\cite[Chs.\@ \textit{A Collection
+of Handy General-Purpose Tactics} and \textit{A Gentle
+Introduction}]{Pierce:SF2}. This includes tactics like \texttt{introv} for
+introducing variables automatically and \texttt{inverts} for performing the
+usually necessary substitutions from \texttt{inversion} on a hypothesis.
+
+We also use Ltac to abbreviate a custom induction principle (see
+\S~\ref{S:coq_ind}) and to automate proofs of the computational well-formed
+property in concrete cases. The custom tactic \texttt{wf\_easy} can prove that
+all of the well-formed examples given in \S~\ref{S:matrix_defn} are indeed
+well-formed. It is particularly useful to solve base cases in proofs. The
+\texttt{wf\_easy} tactic unfolds a few relevant definitions before repeatedly
+applying a strategy to prove the recursive conjuncts generated by the
+computation definition.
+
+\subsubsection{Guiding proof search with hint databases}\label{S:coq_hint}
+
+The \texttt{auto} tactic searches for proof chains in the hypotheses and a core
+database of facts called ``hints.'' We can provide new databases full of
+domain-specific hints; \texttt{auto with hints} will instruct the \texttt{auto}
+tactic to search \texttt{hints} in addition to the hypotheses and the core
+database.
+
+We create a hint database \texttt{matrix} and add the relevant constructors,
+definitions, and theorems to it as they are defined and proved. This allows
+subsequent proofs to benefit from \texttt{auto with matrix} to solve goals that
+use properties of the matrices. In particular, this helps us focus on the
+conceptual challenges and connections of the proof rather than on the details of
+searching for the right chain of hypotheses and theorems. Because this hint
+database is exported, consumers of the library can similarly benefit.
+
+\subsubsection{Enhancing induction through proof computation of induction
+principles}\label{S:coq_ind}
+
+As we learned from the chapter \textit{Induction Principles} of \textit{Logical
+Foundations}~\cite{Pierce:SF1}, Coq permits us to develop custom induction
+principles that can change the way induction works. For example, it may be
+useful to do induction on the length of a list or skipping by twos in the
+naturals, rather than the standard induction on the list itself or by only
+successor in the naturals. As long as we can provide a function with the correct
+form that computes a proof of a general proposition \(P\) over values of the
+type with certain assumed hypotheses (the base and inductive cases), we can use
+\texttt{induction x using IP} to perform induction on \(x\) with
+induction principle \texttt{IP}.
+
+In our case, the induction principle for the type \(\func{matrix\_content}\) was
+not powerful enough---the apparently recursive definition of \(\func{Matrix}\),
+which contains a list of \(\func{matrix\_content}\)s, does not come with an
+induction hypothesis that all the elements of the list obey the property under
+induction. However, we can define an induction principle (called
+\texttt{matrix\_content\_ind\_strong} in our development) which, given a proof
+for the scalar base case and a proof for the recursive matrix case \emph{under
+the assumption that all the elements of the list also satisfy the property},
+correctly computes \(\forall A, (m: \func{matrix\_content}\, A) : P\, m\). It does
+so using list induction under the hood.
+
+We provide the \texttt{mc\_ind} tactic as shorthand for induction using the
+above principle. We use it in most proofs by induction on the contents of a
+matrix. Because the induction hypothesis is strengthened, however, we often have
+to make it more general (\textit{e.g.}, by generalizing the shape of the matrix)
+in order to successfully complete the proof.
+
 \section{Old / Superceded}

### 2021-05-03 paper/paper.tex
paper: bring over justin's work & copy-edit

@@
 but only a subset are used in our program. To illustrate the operations used, we
-begin with a visualization of a three-dimensional array \(E\) with dimensions
-\(3 \times 4 \times 2\) below and discuss how it is used in \matlab.
+begin with the following visualization of a three-dimensional array \(E\) with
+dimensions \(3 \times 4 \times 2\) and discuss how it is used in \matlab.
 
@@
 
+We made an effort to specify each of the statements and expressions according to
+Backus-Naur form, considering the syntax of which each statement or expression
+was composed and instances of overloading. Our next goal was to model the state
+of our program's core matrix data structure so that we could bind syntax to
+semantics. For reasons that we will soon delve into, we did not progress further
+than that step, so our discussion on syntax and semantics ends here. However,
+had we gotten to the point of tying it all together, it would have looked very
+similar to the \textit{Simple Imperative Programs} chapter in \textit{Logical
+Foundations}~\cite{Pierce:SF1}.
+
+\subsection{Blocked by Built-ins}
+
+Thinking that there was no need to re-invent the wheel, we began our journey in
+defining matrix behavior with an exploration into a standard Coq library. Along
+the way, we discovered a number of interesting implementations and proof
+paradigms. Ultimately, we concluded that these would be too difficult to work
+with and that we were better off developing our own implementation. Here we
+discuss the structures from the Coq library that we considered using and the
+limitations that led us away from them.
+
+\subsubsection{Vector}
+
+The natural candidate in the Coq library to use for representing
+multi-dimensional matrices was the \textsf{Vector} type, which was originally
+implemented in the \colorlib\ library~\cite{BLANQUI_2011} and later ported to
+the Coq standard library. We thought that this would be a good abstraction since
+multi-dimensional matrices can have arbitrarily many dimensions and
+\textsf{Vector}s are polymorphic, allowing them to contain other
+\textsf{Vector}s (provided that all contained vectors are of the same type). To
+aid this discussion, the definition of Coq's \textsf{Vector} is stated below:
+
+\begin{verbatim}
+Inductive t A : nat -> Type :=
+| nil : t A 0
+| cons : forall (h:A) (n:nat),
+           t A n -> t A (S n).
+\end{verbatim}
+
+The key thing to make note of is that \textsf{Vector} is a dependent
+type~\cite{Bove2009,Thorsten_2010}; the number of elements in a \textsf{Vector}
+is part of its type, written \texttt{Vector.t A n} for a vector containing \(n\)
+elements of type \(A\)\footnote{The Coq programmer works with dependent types
+all the time, but may not realize it: hypotheses whose types depend on
+quantified variables (\textit{e.g.}, the hypothesis \(H: n \le 0\)) are
+dependently typed as well.}. Since the length of a particular \textsf{Vector} is
+encoded in its type, it is possible to enforce the requirement that all elements
+in a particular dimension have the same length. From here, we devised a scheme
+to stack \textsf{Vector}s to form, \emph{e.g.}, \texttt{Vector.t (Vector.t A m)
+n} such that this we could create matrices of arbitrarily many dimensions, each
+of which has a fixed number of elements. To support this, we defined a
+type-constructor \textsf{matrix} taking a list of dimensions, which represents a
+\textsf{Vector} of \textsf{Vector}s. It can be used as the type of an arbitrary
+\(n\)-dimensional matrix containing elements of type \(A\). This definition is
+as follows:
+
+\begin{verbatim}
+Fixpoint matrix
+        (A: Type) (dims: list nat) :=
+  match dims with
+  | [] => A
+  | head::tail =>
+        Vector.t (matrix A tail) head
+  end.
+
+\end{verbatim}
+
+This definition worked out nicely at first, and we were able to implement
+behavior such as the linearization and element access of a matrix. Although
+possible, this task was non-trivial due to the arbitrary dimensionality of our
+matrices and due to the proof burden imposed on functions over dependent
+types~\cite{SO_2021_1}. Additionally, the existing implementation of element
+access in Coq's \textsf{Vector} type uses the unfamiliar \textsf{Fin} type, an
+abstraction for carrying proofs that a particular number belongs to finite range
+of numbers. In order to access an element of a \textsf{Vector}, one must supply
+such a proof; for our purposes, we use \texttt{Fin.of\_nat}, which gives either
+such a proof or a witness that the number is outside the range. The complexity
+of using the Coq built-ins was quickly growing, and we could deal with it to a
+point but soon ran into trouble while implementing ranged expressions.
+
+\subsubsection{Out of Options}
+
+The problem we experienced while implementing ranged expressions is complex, so
+here we provide an attempt to explain the essence of the problem. The full
+details are available on StackOverflow, where we asked for
+assistance~\cite{SO_2021_2}. The general flavor of the problem is that when
+accessing indices over a range in a matrix, it is impossible to know \textit{a
+priori} the shape of the returned matrix, which in turn informs the returned
+matrix's type (since it is dependently typed based on its dimensionality). This
+computation can be made dynamically for valid ranges, but we run into issues
+when handling degenerate cases---for instance, a range containing values that
+are out of bounds or a list of indices containing the wrong number of indices
+for a particular matrix. In these situations, it is desirable to not return a
+matrix type at all, since the behavior is undefined and will not result in a
+legitimate matrix. The crux of the problem is that with the dependent types, it
+was difficult to specify what dimensions an invalid matrix should have, and the
+general strategy of using \textsf{Option}'s \textsf{Some} and \textsf{None}
+wrappers also did not work because it requires type specification (which is what
+we were having difficulty doing in the first place). Presented with this
+dilemma, we heeded the advice of a commenter on our StackOverflow post, which
+was to eschew dependent types altogether and build our own implementation of
+matrices. The results of doing so encompass the main body of our work, and are
+described in detail in \S~\ref{S:results}.
+
 \section{Results}\label{S:results}
@@
 
-\section{Old / Superceded}
-
-\subsection{Data Structures and Functions}
-
-Our program is used for image processing; thus, many computations are made on
-images, which are represented as arrays in \matlab\@. Additionally, we deal with
-matrices of images, so we need to model multi-dimensional matrices. Since
-multi-dimensional matrices can have arbitrarily many dimensions, we decided the
-best abstraction for this in Coq would be built on top of the \textsf{Vector}
-type which was originally implemented in the \colorlib\
-library~\cite{BLANQUI_2011} and eventually ported to the Coq standard library.
-In Coq, \textsf{Vector} is a dependent type~\cite{Bove2009,Thorsten_2010}; the
-number of elements in a \textsf{Vector} is part of its type, written
-\texttt{Vector.t A n} for a vector containing \(n\) elements of type \(A\).
-Since \textsf{Vector}s are polymorphic, we noted that a \textsf{Vector} could
-contain \textsf{Vector}s provided they are all the same type (which includes
-length, as noted). That is, we can stack \textsf{Vector}s to form, \emph{e.g.},
-\texttt{Vector.t (Vector.t A m) n}. From there, we defined a type-constructor
-\textsf{Matrix} taking a list of dimensions, which represents a \textsf{Vector}
-of \textsf{Vector}s. It can be used as the type of an arbitrary
-\(n\)-dimensional matrix containing elements of type \(A\).
-
-Due to the implementation of \textsf{Vector} in the Coq standard library, we
-needed to greatly extend the functionality of our \textsf{Matrix} type. To that
-end, we have begun implementing \textsf{Matrix} functions such as element
-access, which is non-trivial for matrices with an arbitrary number of
-dimensions\footnote{In particular because of the dependent types; see also
-\url{https://stackoverflow.com/q/66863226/4400820}}. Additionally, the existing
-implementation of element access in Coq's \textsf{Vector} type uses the
-unfamiliar \textsf{Fin} type, an abstraction for carrying proofs that a
-particular number belongs to finite range of numbers. In order to access an
-element of a \textsf{Vector}, one must supply such a proof; for our purposes, we
-use \texttt{Fin.of\_nat}, which gives either such a proof or a witness that the
-number is outside the range. This behavior is captured by the \textsf{sumbool}
-built-in, which required further research on our behalf. Given the number of
-ways that matrices can be indexed in \matlab\ and the complexity introduced by
-Coq's implementation of \textsf{Vector}s, we see this as the main bottleneck for
-our progress. We still need to implement many \textsf{Matrix} operations, such
-as accesses over ranges and over an entire dimension.
-
-\subsection{Binding Operations and Semantics}
-
-While we have not yet bound any of our functions or data types to the statements
-or expressions in \S~\ref{S:mmatlab}, we have naturally been thinking about them
-throughout the whole process. The functions we write are intended to implement
-the evaluation of the corresponding expressions. In the style of
-\texttt{Imp}~\cite{Pierce:SF1}, we intend to write a deterministic expression
-evaluation function; we will then provide semantics for program execution.
-Execution cannot be defined as a function in Coq because it may be
-non-terminating. We have had to consider what operations are be allowed and
-whether checks for undesired behavior (\emph{e.g.}, dynamic type mismatches)
-should be implemented in the underlying functions or in a semantic check.
-
-\section{Goals and Timeline}
-
-As we neglected to include them in our original proposal, we now propose the
-following goals, taking into consideration the progress that we have made so far
-with the project:
-
-\begin{enumerate}
-
-    \item \textbf{Minimum Viable Product}: Successful modelling of a majority of
-        the desired features corresponding to matrix operations in \mmatlab\@.
-        Given the lack of multi-dimensional matrix libraries in Coq and the
-        necessary additions of several features in the Coq standard library,
-        this alone would be a notable achievement for us.
-
-    \item \textbf{Desired Goal}: Successful binding of our matrix operations to
-        \mmatlab\ statements and expressions, including the ability to evaluate
-        programs written in \mmatlab\@. A proof of termination, or sketch of
-        one, for our V1 algorithm is provided.
-
-    \item \textbf{Stretch Goal}: All of the above, and the addition of all
-        operations used in V2 and V3 algorithms (most notably, for loops that
-        increment across a range at varying intervals). Proofs that use
-        these features.
-
-\end{enumerate}
-
-Given what we have encountered so far, we propose the following modified
-timeline for project completion.
-
-\begin{itemize}
-    \item \textbf{03/31/21}: Progress Report Due
-    \item \textbf{04/09/21}: Significant progress made on most matrix operations
-        and data structures.
-    \item \textbf{04/16/21}: Binding \mmatlab\ syntax to operations written
-        in Coq; definition of semantics and program evaluator; formal proof
-        specification written.
-    \item \textbf{04/24/21}: Proof(s) should be written or scope adjusted
-        accordingly. All logic should be implemented; edits for readability
-        should be the main changes henceforth.
-    \item \textbf{04/26/21}: Final Proof Due
-    \item \textbf{05/03/21}: Final Paper Due
-    \item \textbf{05/05/21}: Final Presentation Delivered
-\end{itemize}
-
-We provide the following ``diff'' between the scope of our original project
-proposal and our new proposal:
-\begin{itemize}
-    \item Heightened emphasis on implementation of matrix operations, since they
-        are proving to be non-trivial.
-    \item Addition of goal tiers.
-    \item Discussion on Hoare logic.
-\end{itemize}
-
 {\printbibliography}

### 2021-05-03 paper/paper.tex
paper: abstract

@@
 \begin{abstract}
-    We report on a study of previous set of algorithms that we developed for
-    building mosaics, implemented in \matlab\@. In particular, we are interested
-    in the formalization of the algorithms and the relevant \matlab\ structures
-    in Coq~\cite{Coq} and the termination of these algorithms under appropriate
-    to-be-determined conditions. We present details on the algorithms along with
-    our motivations for studying them. We then discuss an initial foray into
-    related work. Next, we show an updated timeline for our progress. Lastly, we
-    briefly identify the major contributions of each author to this proposal.
+    We report on the development developing matrix algorithms in Coq intended to
+    serve proofs about a subset of \matlab\@. Specifically, we are interested in
+    a previous set of algorithms that we developed for building mosaics,
+    implemented in \matlab\@. We discuss the formalization of the necessary
+    matrix algorithms and the relevant \matlab\ structures in Coq~\cite{Coq}, as
+    well as what remains to be done to prove the termination of the mosaic
+    algorithms under appropriate to-be-determined conditions. We present details
+    on the matrix development along with the key theorems for its correctness.
+    We describe the challenges we faced in developing the matrix type and its
+    operations, providing a glimpse at dependently-typed vectors in Coq. We also
+    briefly discuss advanced automation in Coq useful for the matrix
+    development.
 \end{abstract}

### 2021-05-03 paper/paper.tex
paper: intro/related work

@@
 
-\section{Summary of Proposal}
+\section{Introduction}
 
@@
 source and prove, if possible, their termination given well-conditioned input.
-With some of the work now behind us, we have a better idea of what this project
-entails and how we will be able to reach the finish line.
-
-\section{Problem Description}
+Thus we embarked on developing a Coq model of the mosaic algorithms in order to
+write proofs about them. As we will explain, we did not complete this
+objective---instead, we made progress on the core of the problem: defining
+matrices and their operations. To that extent, we will gloss over the finer
+details of \matlab\ where possible and focus specifically on the relevant matrix
+operations.
+
+In the subsequent sections, we cover related work (\S~\ref{S:related_work}),
+describe the problem we tackled (\S~\ref{S:problem_desc}), our approach to
+designing a solution (\S~\ref{S:design}), and our results (\S~\ref{S:results}).
+
+\section{Related Work}\label{S:related_work}
+
+Because we are proving something about code that we have written ourselves,
+there is not a vast body of literature specific to our algorithms from which we
+can draw. There are however a number of related issues exposed by our code,
+which we will discuss here. In particular, the following are of interest and
+have been written about:
+
+\begin{itemize}
+    \item The correctness of \matlab's \textit{cell array} data structure and
+        its representation in Coq;
+    \item The correctness of \matlab's built-in numerical functions and matrix
+        operations; and
+    \item Proving termination of a particular program based on the inputs of the
+        program.
+\end{itemize}
+
+We are not the first group to demonstrate interest in proving properties about
+matrix representations. In the early days of Coq, a
+paper~\cite{magaud:hal-00955444} was published discussing the implementation of
+square matrices in Coq, as well as proofs involving basic operations such as
+addition and multiplication. This paper references \textit{Linear
+Algebra}~\cite{lin_alg}, a Coq library written by a colleague which formalizes a
+well-known linear algebra text~\cite{linear_algebra} and supports matrices with
+a differing number of rows and columns. These implementations are specific to
+two-dimesional matrices, as far as we can tell, and are thus inadequate for the
+manipulation of higher dimensions in our algorithms.
+
+\matlab\ is a closed-source piece of software, and little has been published about
+its correctness. In the official \matlab\ forum, users have demonstrated interest
+in verifying the correctness of \matlab\ algorithms~\cite{duenisch_2013} and the
+\matlab\ staff reassure users that the algorithms have gone through rigorous
+testing, but this assurance does not carry the same weight as a formal
+verification. Indeed, as we have learned, these techniques fall on separate
+parts of a correctness spectrum, with testing demonstrating correctness less
+reliably than formal methods. The lack of related work in the \matlab\ domain
+served as further motivation to obtain results about the correctness of the subset
+of \matlab\ features used in our code.
+
+Program termination is a topic of wide interest in the computing community, as
+it has both practical and of theoretical relevance given the famous halting
+problem. \citeauthor{Cook_2011}~\cite{Cook_2011} argue in their article that
+computer scientists are often hesitant to make claims about the termination of a
+particular program, but that it is often achievable. Their work draws from the
+field of compilers and makes the argument that one can prove the termination of
+a program given certain inputs, provided that there is sufficient analysis of
+the relation between the inputs and the possible paths that may be taken during
+execution\footnote{In many contexts, this involves things like symbolic
+execution (cf. Rosette~\cite{Rosette}).}. Many of the ideas in this work are
+formalized by \citeauthor{BLANQUI_2011} in the \colorlib\ library for Coq,
+which is used primarily for proofs involving rewriting code and program
+termination~\cite{BLANQUI_2011}.
+
+\section{Problem Description}\label{S:problem_desc}
 

### 2021-05-03 paper/paper.tex
paper: clarify theorem statement

@@
 \begin{thm}[Computing the shape normalizes it]\label{Th:compute_shape_wf_normalizes}
-    For all matrices \(m\), if \(m\) is well-formed than so is the following
-    matrix: \(\{ \var{shape} = \func{compute\_shape}\, m; \var{contents} =
-    \func{contents}\, m \}\).
+    For all matrices \(m\) with contents \(c\), if \(m\) is well-formed than so
+    is the following matrix: \(\{ \var{shape} = \func{compute\_shape}\, c;
+    \var{contents} = c \}\).
 \end{thm}

### 2021-05-03 paper/paper.tex
Update paper.tex

@@
 \begin{abstract}
-    We report on the development developing matrix algorithms in Coq intended to
-    serve proofs about a subset of \matlab\@. Specifically, we are interested in
+    We report on the development of matrix algorithms in Coq intended to
+    serve as proofs about a subset of \matlab\@. Specifically, we are interested in
     a previous set of algorithms that we developed for building mosaics,
@@
 two-dimensional example. Imagine a general two-dimensional \(3 \times 4\) matrix,
-such as the one below, where:
+such as the one below:
 \begin{equation*}
@@
 which unpacks a matrix for use with \(\func{well\_formed'}\). With this
-definition, we can theorems in terms of matrices that are assumed to be
+definition, we can write theorems in terms of matrices that are assumed to be
 well-formed or that must be shown to be well-formed.
